### 2015-03-18 paper/main.tex
add paper latex.

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[letterpaper,margin=0.65in]{geometry}
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage{amssymb}
\usepackage[normalem]{ulem}
\usepackage{natbib}

\def\aap{{ A\&A}}
\def\AaA{{ A\&A}}
\def\AJ{{AJ}}
\def\aj{{AJ}}
\def\ApJ{{ApJ}}
\def\apj{{ApJ}}
\def\apjl{{ApJL}}
\def\ApJL{{ApJL}}
\def\ApJSupp{{ApJS}}
\def\MN{{MNRAS}}
\def\mnras{{MNRAS}}
\def\pasp{{PASP}}
\def\jcap{JCAP}
\def\pasj{{PASJ}}
\def\nat{Nature}
\def\ARAA{{ARA\&A}}
\def\apjs{{ApJS}}
\def\araa{{ARA\&A}}
\def\apss{{Ap\&SS}}
\def\cjaa{{ChJAA}}
\def\aaps{{A\&AS}}

\newcommand{\eg}{{e.g., }}
\newcommand{\ie}{{i.e., }}

\title{A Hybrid Ensemble Learning Approach to Star-galaxy Classification}

\author[E. J. Kim et al.]{
  Edward J. Kim\thanks{jkim575@illinois.edu}$^1$, Robert J. Brunner$^2$,
  and Matias Carrasco Kind$^3$\\
$^1$Department of Physics, University of Illinois, Urbana, IL 61801 USA\\
$^2$Department of Astronomy, University of Illinois, Urbana, IL 61801 USA\\
$^3$National Center for Supercomputing Applications, Urbana, IL 61801 USA}

\begin{document}

\date{\today}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

\maketitle

\label{firstpage}
\begin{abstract}
\textcolor{red}{Abstract goes here.}
\end{abstract}

\begin{keywords}
\textcolor{red}{keywords}
\end{keywords}

\section{Introduction}
  \label{section:introduction}

The problem of star-galaxy classification is fundamental to astronomy
and goes as far back as \cite{messier1781catalogue}.
A variety of different strategies have been developed 
to tackle this long-standing problem,
and yet it remains an unsolved challenge.
The most commonly used method to classify stars and galaxies
in large sky surveys is the morphological separation
~\citep{sebok1979optimal, kron1980photometry, valdes1982resolution,
yee1991faint, vasconcellos2011decision,
henrion2011bayesian}.
It relies on the assumption that
stars appear as point sources
while galaxies appear as resolved sources.
However,
currently ongoing and upcoming large photometric surveys,
such as the Dark Energy Survey
(DES\footnote{http://www.darkenergysurvey.org/})
and the Large Synoptic Survey Telescope
(LSST\footnote{http://www.lsst.org/lsst/}),
will detect a vast number of unresolved galaxies
at faint magnitudes.
Near a survey's limit, the photometric observations
cannot reliably separate stars from unresolved galaxies
by morphology alone without leading to
incompleteness and contamination in the star and galaxy samples.

The contamination of unresolved galaxies can be mitigated
by using training based algorithms.
Machine learning methods have the advantage that
it is easier to include extra information,
such as concentration indices, shape information,
or different model magnitudes.
However,
they are only reliable within the limits of the training data,
and it can be difficult to extrapolate these algorithms
outside the parameter range of the training data.
These techniques can be further categorized into
supervised and unsupervised learning approaches.

In supervised learning, 
the input attributes (\eg magnitudes or colors)
are provided along with the truth labels (\eg star or galaxy).
\cite{odewahn1992automated} pioneered
the application of neural networks
to the star-galaxy classification problem,
and it has become a core part of
the astronomical image processing software
\textsc{SExtractor}~\citep{bertin1996sextractor}.
Other successfully implemented examples include
decision trees~\citep{weir1995automated, suchkov2005census, ball2006robust}
and Support Vector Machines~\citep{Fadely2012}.
Unsupervised machine learning techniques
are less common
as they do not utilize the truth labels during the training process
and only the input attributes are used.

Physically based template fitting methods
have also been used for the star-galaxy classification
problem~\citep{robin2007stellar, Fadely2012}.
Template fitting approaches infer a source's properties
by finding the best match between
the measured set of magnitudes (or colors)
and the synthetic set of magnitudes (or colors)
computed from a set of spectral templates.
Although it is not necessary to obtain
a high-quality spectroscopic training sample,
these techniques do require
a representative sample of theoretical or empirical templates
that span the possible spectral energy distributions (SEDs)
of stars and galaxies.
Furthermore, they are not exempt from uncertainties
due to measurement errors on the filter response curves,
or from mismatches between the observed magnitudes
and the template SEDs.

In this paper,
we present a novel star-galaxy classification framework
that combines and fully exploits different classification techniques
to produce a more robust classification.
In particular,
we show that the combination of a morphological separation method,
a template fitting technique, a supervised machine learning method,
and an unsupervised machine learning algorithm
can improve the overall performance over any individual method.
In Section~\ref{section:classification_methods},
we describe each of the star-galaxy classification methods.
In Section~\ref{section:classification_combination_methods},
we describe different classification combination techniques
In Section~\ref{section:data},
we describe the Canada-France Hawaii Telescope Lensing Survey (CFHTLenS)
data set with which we test the algorithms.
In Section~\ref{section:results_and_discussion},
we compare the performance of our combination techniques
to the performance of the individual classification techniques.

\section{Classification Methods}
  \label{section:classification_methods}

In this section, we present
four distinct star-galaxy classification techniques.
The first method is
a supervised machine learning technique named \texttt{TPC}
(Trees for Probabilistic Classification),
which uses prediction trees and a random forest~\citep{carrascokind2013tpz}.
The second method is
an unsupervised machine learning technique named \texttt{SOMc},
which uses self-organizing maps (SOM) and a random atlas
to provide a classification~citep{carrascokind2014somz}.
The third method is
a Hierarchical Bayesian template fitting technique
based on the work by \cite{Fadely2012},
which fits SED templates from star and galaxy libraries
to an observed set of measured flux values.
The fourth method is a morphological separation method,
which uses a hard cut in the half-light radius vs.\ magnitude plane.

Collectively, these four methods represent the majority of all
standard star-galaxy classification approaches published in the literature.
It is very likely that any new classification technique would be
functionally similar to one of these four   methods.
Therefore, any of these four methods could in principle be replaced by a similar method.

\subsection{Morphological Separation}

The simplest and perhaps the most widely used approach
to star-galaxy classification is
to make a hard cut in the space of photometric attributes.
As a first-order morphological selection of point sources,
we adopt a technique that is popular among the weak lensing 
community~\citep{Kaiser1995}.
As Figure~\ref{fig:morph} shows, there is a distinct locus
produced by point sources in the half-light radius
($r_h$; estimated by \textsc{SExtractor}'s \texttt{FLUX\_RADIUS} parameter)
vs.\ the $i$-band magnitude plane.
A rectangular cut in this size-magnitude plane separates point sources,
which are presumed to be stars,
from resolved sources,
which are presumed to be galaxies.
The boundaries of the selection box are determined by
manually inspecting the size-magnitude diagram.

\begin{figure}
  \centering
  \includegraphics[width=0.95\linewidth]{figures/morph.eps}
  \caption{Half-light radius ($r_h$) vs.\ magnitude.}
  \label{fig:morph}
\end{figure}

One of the disadvantages of such cut-based methods is
that it classifies every source with absolute certainty.
It is difficult to justify such a decisive classification
near a survey's magnitude limits,
where measurement uncertainties generally increase.
A more informative approach is to provide probabilistic classifications.
Although a recent work by \cite{henrion2011bayesian}
implemented a probabilistic classification using a Bayesian approach
on the morphological measurements alone,
here we use a cut-based morphological separation
to demonstrate the advantages of our combination techniques.
In particular, we later show that
the binary output (\ie 0 or 1) of cut-based methods
can be transformed into a
probability estimate by combining them
with the probability outputs from other
probabilistic classification techniques,
such as \texttt{TPC}, \texttt{SOMc}, and \texttt{HB}.

\subsection{Supervised Machine Learning: TPC}

\texttt{TPC} is a parallel, supervised machine learning algorithm
that uses prediction trees and random forest 
techniques~\citep{breiman1984classification, breiman2001random}
to produce star-galaxy classification.
\texttt{TPC} is a part of a publicly available software package called
\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}
(Machine Learning for Photo-$z$).
The full software package includes, among other capabilities,
\texttt{TPZ}, a supervised photo-$z$ technique
\citep[regression mode;][]{carrascokind2013tpz},
\texttt{TPC}, a supervised star-galaxy classification technique
(classification mode),
\texttt{SOM}$z$, an unsupervised photo-$z$ technique
\citep[regression mode;][]{carrascokind2014somz},
and
\texttt{SOMc}, an unsupervised star-galaxy classification technique
(classification mode).

\texttt{TPC} uses classification trees,
a type of prediction trees that are designed to
provide a classification or predict a discrete category.
Prediction trees are built by asking a sequence of questions
that recursively split the data into branches
until a terminal leaf is created
that meets a stopping criterion
(\eg a minimum leaf size).
The optimal split dimension is decided by
choosing the attribute that maximizes
the \textit{Information Gain} ($I_G$), which is defined as

\begin{equation}
  I_G \left(D_{\mathrm{node}}, X\right)
  = I_d \left( D_{\mathrm{node}} \right)
  - \sum_{x \in \mathrm{values}(X)}
  \frac{|D_{\mathrm{node}, x}|}{|D_{\mathrm{node}}|}
  I_d \left( D_{\mathrm{node}, x} \right),
  \label{eq:information_gain}
\end{equation}

\noindent
where $D_{\mathrm{node}}$ is the training data in a given node,
$X$ is one of the possible dimensions (\eg magnitudes or colors)
along which the node is split, and
$x$ are the possible values of a specific dimension $X$.
$|D_{\mathrm{node}}|$ and $|D_{\mathrm{node}, x}|$ are the size of the total training data
and the number of objects in a given subset $x$ within the current node,
respectively.
$I_d$ is the impurity degree index,
and \texttt{TPC} can calculate $I_d$
from any of the three standard different impurity indices:
\textit{information entropy}, \textit{Gini impurity},
and \textit{classification error}.
Here, we use the information entropy,
which is defined similarly to the thermodynamic entropy:

\begin{equation}
  I_d \left( D \right)
  = - f_g \log_{2} f_g - (1 - f_g) \log_{2} f_g,
\end{equation}

\noindent
where $f_g$ is the fraction of galaxies in the training data.
At each node in our tree,
we scan all dimensions to identify the split point that
maximizes the information gain as defined by Equation~\ref{eq:information_gain},
and select the attribute that maximizes the impurity index overall.

In a technique called random forest,
we create bootstrap samples
(\ie $N$ randomly selected objects with replacement)
from the input training data
by sampling repeatedly from the magnitudes using the magnitude errors.
We use these bootstrap samples to construct
multiple, uncorrelated prediction trees
whose individual predictions are aggregated to produce
a star-galaxy classification for each source.

We also use a cross validation technique called
Out-of-Bag~\citep[OOB;][]{breiman1984classification, carrascokind2013tpz}.
When a tree (or a map) is built in \texttt{TPC} (or \texttt{SOMc}),
a fraction of the training data, usually one-third,
is left out and not used in training the trees or maps.
After a tree is constructed using two-thirds of the training data,
the final tree is applied to the remaining one-third
to make a classification.
This process is repeated for every tree,
and the predictions from each tree are aggregated
for each object to make the final star-galaxy classification.
We emphasize that if an object is used for training a given tree,
it is never used for subsequent prediction by that tree.
Thus, the OOB data is an unbiased estimation of the errors
and can be used as cross-validation data
as long as the OOB data remain similar to the final test data set.
The OOB technique can also provide extra information such as
a ranking of the relative importance of the input attributes
used in the prediction.
The OOB technique can prove extremely valuable
when calibrating the algorithm,
when deciding which attributes to incorporate in the construction of the trees,
and when combining this approach with other techniques.

\subsection{Unsupervised Machine Learning: SOMc}

A Self-Organizing Map
\citep[SOM;][]{kohonen1990self, kohonen2001self}
is an unsupervised, artificial neural network algorithm
that is capable of projecting high-dimensional input data
onto a low-dimensional map
through a process of competitive learning.
In astronomical applications,
the high-dimensional input data can be
magnitudes, colors, or some other photometric attributes.
The output map is usually chosen to be two-dimensional
for the visualization.
The differences between a SOM and other neural network algorithms are
that a SOM is unsupervised,
there are no hidden layers and therefore no extra parameters,
and it produces a direct mapping
between the training set and the output network.
In fact, a SOM can be viewed as a non-linear generalization
of a principal component analysis (PCA) algorithm.

The key characteristic of SOM is that
it retains the topology of the input training set,
revealing correlations between input data that are not obvious.
The method is unsupervised:
the user is not required to specify the desired output
during the creation of the lower-dimensional map,
and the mapping of the components from the input vectors
is a natural outcome of the competitive learning process.

During the construction of a SOM,
each node on the two-dimensional map is represented by
weight vectors of the same dimension
as the number of attributes used to create the map itself.
In an iterative process,
each object in the input sample is individually used
to correct these weight vectors.
This correction is determined so that the specific neuron (or node),
which at a given moment best represents the input source,
is modified along with the weight vectors
of that node's neighboring neurons.
As a result, this sector within the map
becomes a better representation of the current input object.
This process is repeated for every object in the training data,
and the entire process is repeated for several iterations.
Eventually, the SOM converges to its final form where
the training data is separated into groups of similar features.

In a similar approach to random forest in \texttt{TPZ} and \texttt{TPC},
\texttt{SOM}$z$ uses a technique called \textit{random atlas}
to provide photo-$z$ estimation \citep{carrascokind2014somz}.
In random atlas, the prediction trees of random forest are replaced by maps,
and each map is constructed from different bootstrap samples
of the training data.
Furthermore, we create random realizations of the training data
by perturbing the input attributes by their measurement error.
For each map, we can either use all available attributes,
or randomly select a subsample of the attribute space.
This SOM implementation can also be
applied to the classification problem,
and we refer to it as \texttt{SOMc}
in order to differentiate it from 
the photo-$z$ estimation problem (regression mode).
We also use the random atlas approach
in some of the classification combination approaches as discussed in
Section~\ref{section:classification_combination_methods}.

One of the most important parameter in \texttt{SOMc}
is the topology of the two-dimensional SOM,
which can be rectangular, hexagonal, or spherical.
To classify stars and galaxies in the CFHTLenS data,
we use a spherical topology,
which is constructed by using
\texttt{HEALPIX}~\citep{gorski2005healpix}.
Furthermore, similar to \texttt{TPC},
we use the OOB technique to make an unbiased estimation of errors.
For a complete description of the SOM implementation
and its application to the estimation of
photo-$z$ probability distribution funtions,
we refer the reader to \cite{carrascokind2014somz}.

\subsection{Template fitting: Hierarchical Bayesian}

One of the most common methods to classify a source
based on its observed magnitudes is template fitting.
Template fitting algorithms are unsupervised;
there is no need for additional knowledge outside the
observed data and the template SEDs.
However, any incompleteness in our knowledge of the template SEDs
that fully span the possible SEDs of observed sources
may lead to misclassification of sources.

Bayesian algorithms use Bayesian inference to quantify
the relative probability that each template matches
the input photometry
and determines a probability estimate by computing
the posterior that a source is a star or a galaxy.
In this work, we have modified and parallelized
a publicly available Hierarchical Bayesian (HB) template fitting
algorithm by \cite{Fadely2012}.
In this section, we provide a brief description
of the HB template fitting technique;
for the details of the underlying HB approach,
we refer the reader to \cite{Fadely2012}.

We write the posterior probability that a source is a star as

\begin{equation}
P \left( S | \mathbf{x}, \mathbf{\theta} \right)
= P \left( \mathbf{x} | S, \mathbf{\theta} \right)
P \left( S | \mathbf{\theta} \right),
\end{equation}

\noindent
where $\mathbf{x}$ represents a given set of observed magnitudes,.
We have also introduced the \textit{hyperparameter} $\mathbf{\theta}$,
a nuisance parameter that characterizes our uncertainty
in the prior distribution.
To compute the likelihood that a source is a star,
we marginalize over all star and galaxy templates $\mathbf{T}$.
In a template-fitting approach,
we marginalize by summing up
the likelihood that a source has the set of magnitudes $\mathbf{x}$
for a given star template
as well as the likelihood for a given galaxy template:

\begin{equation}
  P \left(\mathbf{x} | S, \mathbf{\theta} \right)
  = \sum_{t \in \mathbf{T}}
  P \left(\mathbf{x} | S, t, \mathbf{\theta} \right)
  P \left(t | S, \mathbf{\theta} \right).
  \label{eq:marginalize_template}
\end{equation}

\noindent
The likelihood of each template
$P \left( \mathbf{x} | S, \mathbf{\theta} \right)$
is itself marginalized over the uncertainty
in the template-fitting coefficient.
Furthermore, for galaxy templates, we introduce another step that 
marginalizes the likelihood by redshifting a given galaxy template
by a factor of $1 + z$.

Marginalization in Equation~\ref{eq:marginalize_template}
requires that we specify the prior probability
$P \left(t | S, \mathbf{\theta} \right)$
that a source has spectral template $t$ (at a given redshift).
Thus, the probability that a source is a star (or a galaxy)
is either the posterior probability itself
if a prior is used,
or the likelihood itself if an uniformative prior is used.
In a Bayesian analysis, it is preferable to use a prior,
which can be directly computed either from physical assumptions,
from an empirical function calibrated by
using a spectroscopic training sample,
or from an empirical function calibrated by
using machine learning techniques.
In an HB approach, the entire sample of sources is used to
infer the prior probabilities for each individual source.

Since the templates are discrete in both SED shape and physical properties,
we parametrize the prior probability of each template
as a discrete set of weights such that

\begin{equation}
\sum_{t \in \mathbf{T}}
P \left(t | S, \mathbf{\theta} \right) = 1.
\end{equation}

\noindent
These weights correspond to the hyperparameters,
which can be inferred by sampling
the posterior probability distribution in the hyperparameter space.
For the sampling, we use \texttt{emcee}, a Python implementation of the
affine-invariant Markov Chain Monte Carlo (MCMC) ensemble sampler~\citep{Foreman-Mackey2013}.

As the goal of template fitting methods is to minimize
the difference between observed and theoretical magnitudes,
this approach heavily relies on
both the use of SED templates and
the accuracy of the transmission functions
for the filters used for particular survey.
For our stellar templates,
we use the empirical SED library from \cite{pickles1998stellar}.
The Pickles library consists of 131 stellar templates,
which span all normal spectral types
and luminosity classes at solar abundance,
as well as metal-poor and metal-rich F-K dwarf 
and G-K giant and supergiant stars.
For our galaxy templates,
we use the four CWW spectra from \cite{coleman1980colors},
which include an Elliptical, an Sba, an Sbb,
and an Irregular galaxy template.
When extending an analysis to higher redshift,
the CWW library is often augmented with
two star bursting galaxy templates from \cite{kinney1996template}.
\textcolor{red}{interpolated}.

All of the above templates are convolved
with the filter response curves to generate model magnitudes.
These response curves consist of
$u$, $g$, $r$, $i$, $z$ filter transmission functions
for the observations taken by the
Canada-France Hawaii Telescope (CFHT).

\section{Classification combination methods}
  \label{section:classification_combination_methods}

Building on the work in the field of ensemble learning,
we combine the predictions from
individual star-galaxy classification techniques
using ensemble learning techniques known as
Bayesian Model Combination (BMC)
and a variant of an ensemble learning technique known as stacking.
The main idea behind ensemble learning is to weigh
the predictions from individual models
and combine them to obtain a prediction
that outperforms every one of 
them individually~\citep{rokach2010ensemble}.

\subsection{Unsupervised Binning}
  \label{section:random_atlas}

Given the variety of star-galaxy classification methods
we are using,
we fully expect the relative performance
of the individual techniques to vary across
the parameter space spanned by the data.
For example, it is reasonable to expect 
supervised techniques to outperform other techniques
in areas of parameter space that are well-populated
with training data.
Similarly, we can expect unsupervised approaches
such as \texttt{SOM} or template fitting approaches 
to generally perform better when training data
is either sparse or unavailable.

We therefore adopt the binning strategy similar to 
\cite{carrascokind2014exhausting}.
In this binning strategy,
we allow different classifier combinations 
in different parts of parameter space
by creating two-dimensional SOM representations of
the full nine-dimensional magnitude-color space.
A SOM representation can be rectangular, hexagonal, or spherical;
here we choose a 10$\times$10 rectangular topology to facilitate 
visualization as shown in Figure.~\ref{fig:som_colors}.
For all combination methods,
we use only the OOB (cross validation) data contained in each cell
to compute the relative weights for the base classifiers.
The weights within individual cells are then applied to the test data
to make the prediction.

Furthermore, we construct a collection of SOM representations
and subsequently combine the predictions from each map
into a meta-prediction.
Given a training sample of $N$ sources,
we generate $N_R$ random realizations of training data
by perturbing the attributes
with the measured uncertainty for each attribute.
The uncertainties are assumed to be normally distributed.
In this manner,
we reduce the bias towards the data
and introduce randomness in a systematic manner.
For each random realization of a training sample,
we create $N_M$ bootstrap samples of size $N$
to generate $N_M$ different maps.

After all maps are built,
we have a total of $N_R \times N_M$ probabilistic outputs
for each of the $N$ sources.
To produce a single probability estimate for each source,
one can take the mean, the median, or some other simple statistic.
With a sufficient number of maps,
we find that there is usually negligible difference
between taking the mean and taking the median,
and so we use the median in the following sections.
We note that it is also possible to establish confidence intervals
using the distribution of the probability estimates.

\subsection{Weighted Average}

The simplest approach to combine different combination techniques is
to simply add the individual classifications from the base classifiers
and renormalize the sum.
In this case, the final probability is given by
\begin{equation}
P\left(S | \mathbf{x}, \mathbf{M} \right)
= \sum_{i} P \left( S | \mathbf{x}, M_{i} \right),
\end{equation}

\noindent
where $\mathbf{M}$ is the set of models
(\texttt{TPC}, \texttt{SOMc}, \texttt{HB}, and morphological separation
in our work).
We improve on this simple approach
by using the binning strategy to
calculate the weighted average of each bin separately,
rather than globally.

\subsection{Bucket of Models (BoM)}

After the multi-dimensional input data have been binned,
we can use the cross-validation data
to choose the best model within each bin,
and we use only that model within that specific bin
to make predictions for the test data.
We use the mean squared error
(MSE; also known as Brier score~\citep{brier1950verification})
as a classification error metric. We define MSE as

\begin{equation}
  \mathrm{MSE} = \frac{1}{N} \sum^{N - 1}_{i = 0}
  \left( y_i - \hat{y}_i \right)^2,
  \label{eq:mse}
\end{equation}

\noindent
where $\hat{y}_i$ is the actual truth value
(\eg 0 or 1) of the $i^{\text{th}}$ data, 
and $y_{i}$ is the probability prediction made by the models.
Thus, a model with the minimum MSE is chosen in each bin
and is assigned a weight of one,
and zero for all other models.
However, the chosen model is allowed to vary between different bins.

\subsection{Stacking}

Instead of selecting a single model that performs best within each bin,
we can train a learning algorithm to combine the output values of
several other base classifiers in each bin.
An ensemble learning method of using a meta-classifier
to combine lower-level classifiers is known as \textit{stacking}
or \textit{stacked generalization}~\citep{wolpert1992stacked}.
Although any arbitrary algorithm
can theoreticaly be used as a meta-classifier,
a logistic regression or a linear regression is often used
in practice.
In our work,
we use a single-layer multi-response linear regression algorithm,
which often shows the best performance~\citep{breiman1996stacked, ting1999issues}.

\subsection{Bayesian Model Combination}

We also use a model combination technique \citep{Monteith2011} known as
Bayesian Model Combination (BMC).
BMC uses Bayesian principles to generate an ensemble combination of
different classifiers~\citep{Monteith2011}.
The posterior probability that a source is a star is given by

\begin{equation}
  P \left(S | \mathbf{x}, \mathbf{D}, \mathbf{M}, \mathbf{E} \right)
  = \sum_{e \in \mathbf{E}} P \left(S | \mathbf{x}, \mathbf{M}, e \right)
  P \left(e | \mathbf{D} \right),
  \label{eq:p_star_bmc}
\end{equation}

\noindent
where $\mathbf{D}$ is the data set,
and $e$ is an element in the ensemble space $\mathbf{E}$ of possible model combinations.
By Bayes' Theorem, the posterior probability of $e$ given $\mathbf{D}$ is given by

\begin{equation}
  P \left(e | \mathbf{D} \right)
  = \frac{P \left(e \right)}{P \left(\mathbf{D} \right)}
  \prod_{d \in \mathbf{D}} P \left( d | e \right)
  \propto P \left(e\right) \prod_{d \in \mathbf{D}} P \left(d | e \right).
  \label{eq:p_ensemble}
\end{equation}

\noindent
Here, the product of $P\left(d \mid e\right)$ is
over all individual data $d$ in the training data $\mathbf{D}$,
and $P\left(\mathbf{D}\right)$ is merely a normalization factor
and not important.

For binary classifiers whose output is either zero or one
(\eg a cut-based morphological separation),
we assume that each example is corrupted with
an average error rate $\epsilon$.
This means that
$P\left(d|e\right) = 1-\epsilon$ if the combination $e$
correctly predicts class $\hat{y}_i$ for the $i^{\text{th}}$ object,
and $P\left(d|e\right) = \epsilon$ if it predicts an incorrect class.
The average rate $\epsilon$ can be estimated by
the fraction $\left(M_g + M_s\right) / N$,
where $M_g$ are $M_s$ are defined in Table~\ref{table:def_numbers},
and $N$ is the total number of sources.
Equation~\ref{eq:p_ensemble} then becomes

\begin{equation}
  P \left( e | \mathbf{D} \right) \propto 
  P \left( e \right) \left(1 - \epsilon \right)^{N_s + N_g}
  \left( \epsilon \right)^{M_s + M_g}.
\end{equation}

\noindent
For probabilistic classifiers,
we can directly use the probabilistic predictions
and write Equation~\ref{eq:p_ensemble} as

\begin{equation}
  P \left( e | \mathbf{D} \right) \propto 
  P \left( e \right) \prod_{i=0}^{N-1}
  \hat{y}_i y_i + 
  \left(1 - \hat{y}_i\right) \left(1 - y_i\right).
\end{equation}

\begin{table}
  \caption{The definition of the number of stars/galaxies
           classified and misclassified as stars/galaxies.}
  \centering
  \begin{tabular}{ l | c | c }
    \hline
                           & True galaxies & True stars \\ \hline
    Classified as galaxies & $N_g$ & $M_s$ \\ \hline
    Classified as stars & $M_g$ & $N_s$ \\
    \hline
  \end{tabular}
  \label{table:def_numbers}
\end{table}

Although the space $\mathbf{E}$ of potential model combinations
is in principle infinite,
we can produce a reasonable finite set
of potential model combinations by using sampling techniques.
In our implementation,
the weights of each combination of the base classifiers
is obtained by sampling from a Dirichlet distribution.
We first set all alpha values of a Dirichlet distribution.
We then sample this distribution $q$ times
to obtain $q$ sets of weights.
For each combination,
we assume a uniform prior and
calculate $P\left(e|\mathbf{D}\right)$ using Equation~\ref{eq:p_ensemble}.
We select the combination with the highest $P\left(e|\mathbf{D}\right)$,
and update the alpha values by
adding the weights of the most probable combination
to the current alpha values.
The next $q$ weight sets of weights are drawn
using the updated alpha values.

We continue the sampling process until
we reach a predefined number of combinations,
and finally use Equation~\ref{eq:p_star_bmc} to compute
the posterior probability that a source is a star (or a galaxy).
In this paper, we use a $q$ value of three,
and 1,000 model combinations were considered.

We also use a binned version of the BMC technique,
where we use a SOM representation
to apply different model combinations
for different regions of the parameter space.
We however note that introducing randomness
though the construction of different SOM representations
does not show significant
improvement over using only one single SOM representation.
This similarity is likely due to the randomness 
that has already been introduced by 
the sampling from the Dirichlet distribution.
Thus, our BMC technique uses one SOM,
while other base models (WA, BoM, and stacking)
generate $N_R$ random realizations of $N_M$ maps.

\section{Data}
  \label{section:data}

We use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{Heymans2012,Erben2013,Hildebrandt2012}.
This catalog consits of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degress
in the four fields of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

We have cross-matched reliable spectroscopic galaxies from
the Deep Extragalactic Evolutionary Probe Phase 2~
\citep[DEEP2;][]{Davis2003,Newman2013a},
the Sloan Digital Sky Survey Data Release 10~\citep[SDSS-DR10]{Ahn2014},
the VIsible imaging Multi-Object Spectrograph (VIMOS)
Very Large Telescope (VLT) Deep Survey~
\citep[VVDS;][]{LeFevre2005,Garilli2008}, and
the VIMOS Public Extragalactic Redshift Survey~\citep[VIPERS;][]{Garilli2014},
In the end, we have 8,545 stars and 57,843 galaxies available
for the training and testing processes.

Our goal here is not to obtain the best classifier
performance; for this we would have fine tuned individual base
classifiers and chosen sophisticated models best suited to the
particular properties of the CFHTLenS data.
For example, \cite{Hildebrandt2012} suggest that
all objects with $i > 23$ in the CFHTLenS dataset
may be classified as galaxies
without significant incompleteness in the galaxy sample.
Although this approach works because
the high Galactic latitude fields of the CFHTLS contain
relatively few stars,
it is very unlikely that such an approach
will meet the science requirements for the quality of
star-galaxy classification in lower latitude, star-crowded fields.
Rather, our goal for the CFHTLenS dataset
is to demonstrate the usefulness
of combining different classifiers even when
the base classifiers may be poor or trained on partial data.

\section{Results and Discussion}
  \label{section:results_and_discussion}

In this section, we present the classification performance of
the four different combination techniques,
as well as the individual star-galaxy classification techniques
on the CFHTLenS test data.

\subsection{Classification Metrics}

Probabilistic classification models can be considered as
functions that output a probability estimate of each source
to be in one of the classes (\eg a star or a galaxy).
Although the probability estimate can be used as a weight 
in subsequent analyses to improve or enhance
a particular measurement~\citep{ross2011ameliorating},
it can also be converted into a class label
by using a threshold (probability cut).
The simplest way to choose the threshold is to set it to a fixed value,
\eg $p_\mathrm{cut} = 0.5$.
This is, in fact, what is often done
\citep[\eg][]{henrion2011bayesian, Fadely2012}.
However, choosing $0.5$ as a threshold is not the best choice
for an unbalanced data set, where galaxies outnumber stars.
Furthermore, setting a fixed threshold ignores the operating condition 
(\eg science requirements, stellar distribution, misclassification costs)
where the model will be applied.

When we have no information about the operating condition
when evaluating the performance of classifiers,
there are effective tools such as
the Receiver Operating Characteristic (ROC) curve
\citep{swets2000better}.
An ROC curve is a graphical plot that illustrates the true positive rate
versus the false positive rate of a binary classifier
as its classification threshold (probability cut) is varied.
The Area Under the Curve (AUC) summarizes the curve information
in a single number,
and can be used as an assessment of the overall performance.

In astronomical applications,
the operating condition usually translates to
the completeness and purity requirements of the star/galaxy sample,
We define the galaxy \textit{completeness}
$c_g$ (also known as recall or sensitivity) as
the fraction of the number of true galaxies classified as galaxies
out of the total number of true galaxies,
\begin{equation}
c_g = \frac{N_g}{N_g + M_g},
\end{equation}

\noindent
where $N_g$ and $M_g$ are define in Table~\ref{table:def_numbers}.
We define the galaxy \textit{purity} $p_g$ (also known as precision
or positive predictive value)
as the fraction of the number of true galaxies classified as galaxies
out of the total number of objects classified as galaxies, 
\begin{equation}
p_g = \frac{N_g}{N_g + M_s},
\end{equation}

\noindent
where $N_g$ and $M_g$ are define in Table~\ref{table:def_numbers}.
Star completeness and purity are defined in a similar manner.

One of the advantages of a probabilistic classification is
that the threshold can be adjusted to produce
a more complete but less pure sample,
or a less complete but more pure one.
To compare the performance of probabilistic classification techniques
with that of the morphological separation,
which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
at a certain purity ($p_g = 0.9484$, $p_s = 0.9666$),
we adjust the threshold of probabilistic classifiers
until the galaxy completeness matches that of the morphological classifier
to compute $p_{g}$ at $c_g=0.9964$.
Similarly, the star purity $p_{s}$ at $c_{s}=0.7145$
is computed by adjusting the threshold values
until the star completeness of each classifier is equal to
that of the morphological classifier.

We can also compare the performance of different classification techniques
by assuming an arbitrary operating condition.
For example, weak lensing science measurements
of the DES require $c_g > 0.960$ and $p_g > 0.778$
to control both the statistical and systematic errors
on the cosmological parameters,
and $c_s > 0.250$ and $p_s > 0.970$
for stellar Point Spread Function (PSF) calibration
\citep{soumagnac2013star}.
Although these values will likely be different
for the science cases of the CFHTLenS data,
we adopt these values to compare the classification performance
at a reasonable operating condition.
Thus, we compute $p_{g}$ at $c_g=0.960$
and $c_{s}$ at $p_s=0.250$.
We also use the MSE defined in Equation~\ref{eq:mse}
as a classification error metric.

\begin{table}
  \caption{The definition of the classification performance metrics.}
  \centering
  \begin{tabular}{c l}
  Metric & Meaning \\
  \hline
  AUC & Area under the Receiver Operating Curve \\
  MSE & Mean squared error \\
  $c_g$ & Galaxy completeness \\
  $p_g$ & Galaxy purity \\
  $c_s$ & Star completeness \\
  $p_s$ & Star purity \\
  $p_g(c_g=x)$ & Galaxy purity at $x$ galaxy purity \\
  $p_s(c_s=x)$ & Star purity at $x$ star purity \\
  \end{tabular}
  \label{table:metrics}
\end{table}

\subsection{Classifier Combination}
  \label{section:rich_training}

%% rich training data
\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           as applied to the CFHTLenS data,
           with no cut applied to the training data set.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
%  \begin{tabular}{l c c c c c c}
%    Classifier & AUC & MSE &
%    $p_{g}\left(c_g=0.9964\right)$ & $p_{s}\left(c_s=0.7145\right)$ &
%    $p_{g}\left(c_g=0.9600\right)$ & $p_{s}\left(c_s=0.2500\right)$ \\
%    \hline
%    TPC  & 0.9870 & 0.02081 & 0.9717 & 0.9843 & 0.9925 & 0.9977\\
%    SOMc & 0.9683 & 0.04520 & 0.9099 & 0.8487 & 0.9781 & 0.9844\\
%    HB   & 0.8954 & 0.08777 & 0.9048 & 0.7407 & 0.9593 & 0.7609\\
%    Morphology   & -      & 0.0397 & 0.9597 & 0.9666 & - & -\\
%    WA   & 0.9718 & 0.03292 & 0.9703 & 0.9876 & 0.9766 & 1.0000\\
%    BoM  & 0.9870 & 0.02081 & 0.9733 & 0.9843 & 0.9925 & 0.9977\\
%    Stacking  & \textbf{0.9878} & 0.02043 & 0.9729 & 0.9909 &
%    \textbf{0.9928} & 1.0000\\
%    BMC  & 0.9849 & \textbf{0.01911} & \textbf{0.9797} & \textbf{0.9950} &
%    0.9916 & 1.0000\\
%  \end{tabular}
  \input{tables/metrics_all.tex}
  \label{table:metrics_all}
\end{table*}

We present in Table~\ref{table:metrics_all}
the classification performance obtained by
applying the four different combination techniques,
as well as the individual star-galaxy classification techniques,
on the CFHTLenS test data.
The bold entries highlight the best technique for any particular metric.
The first four rows show the performance of four individual star-galaxy
classification techniques.
Given the wealth of training data, it is not surprising that 
our supervised machine learning tehcnique \texttt{TPC}
outperforms other unsupervised technique.
\texttt{TPC} is thus shown in the first row as the benchmark.

The simplest of the combination techniques, WA and BoM,
generally do not perform better than \texttt{TPC}.
It is also interesting that,
even with binning the parameter space and selecting the best model
within each bin,
BoM almost always chooses \texttt{TPC} as the best model,
and therefore gives the same performance as \texttt{TPC} in the end.
However, our BMC and stacking techniques have a similar performance
and often outperform \texttt{TPC}.
We observe that our stacking algorithm shows the best performance
as measured by the AUC,
while BMC shows the best performance
as measured by the MSE.

\begin{figure*}
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/som_colors.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the mean $i$-band magnitude (top left) and
             the mean values of three colors,
             $u-g$, $g-r$, and $i-z$,
             for the cross validation data.}
    \label{fig:som_colors}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/weights.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the relative weights for the BMC combination technique
             applied to the four individual methods for the CFHTLenS data.}
    \label{fig:weights}
  \end{minipage}
\end{figure*}

In Figure~\ref{fig:som_colors}, we show
the mean CFHTLenS $i$-band magnitude in the top right panel,
and the mean $u-g$, $g-r$, and $i-z$ colors in the remaining three panels.
These two-dimensional maps clearly shows
the ability of the SOM to preserve
relationships between sources
when it projects the full nine-dimensional space to
the two-dimensional map.
The SOM mapping is a non-linear representation of
all magnitudes and colors,
and thus the CHFTLenS $i$-band magnitude and color maps should 
only be used to provide guidance.

We can also use the same SOM to determine the relative weights for 
the four individual classification methods for each cell.
We present the four weight maps for the BMC technique
in Figure~\ref{fig:weights}.
In these maps, a redder color indicates a higher weight,
or equivalently that the corresponding weight performs better in that region.
These weight maps demonstrate the variation in
the performance of the individual techniques across
the two-dimensional parameter space defined by the SOM.
Not surprisingly, the morphological separation
performs best in the lower right corner of the $i$-band magnitude map,
which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
On the other hand, \texttt{TPC} seems to perform best
in the region that corresponds to intermediate magnitudes
$20\gtrsim i \lesssim22.5$ and $1.5<u-g<3.0$.
Our unsupervised learning method \texttt{SOMc}
performs relatively better at fainter magnitudes $i>21.5$
with $0<u-g<0.5$ and $0<g-r<0.5$.
The BMC technique assigns almost no weight to
the worst-performing algorithm, \texttt{HB},
when there is a wealth of training data.

\subsection{Insufficient Training Data}
  \label{section:poor_training}

It is very costly in terms of telescope time to
obtain a large sample spectroscopic observations
down to the limiting magnitude of a photometric sample.
Thus, we investigate the impact of training set quality
by considering a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels from the VVDS
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 6,361 objects.

\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           as applied to the CFHTLenS data in the VVDS field
           with a magnitude cut of $i < 22.0$.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
  \begin{tabular}{l c c c c c c}
    Classifier & AUC & MSE &
    $p_{g}\left(c_g=0.9964\right)$ & $p_{s}\left(c_s=0.7145\right)$ &
    $p_{g}\left(c_g=0.9600\right)$ & $p_{s}\left(c_s=0.2500\right)$\\
    \hline
    TPC  & 0.9601 & 0.0728 & 0.9438 & 0.9329 & 0.9819 & 0.9950\\
    SOMc & 0.9008 & 0.1679 & 0.8843 & 0.5044 & 0.9244 & 0.7726\\
    HB   & 0.8960 & 0.0894 & 0.9074 & 0.7516 & 0.9615 & 0.7793\\
    Morphology   & -      & \textbf{0.0397} & 0.9597 & 0.9666 & - & -\\
    WA   & 0.9639 & 0.0581 & 0.9613 & 0.9787 & 0.9756 & 1.0000\\
    BoM  & 0.9601 & 0.0728 & 0.9438 & 0.9329 & 0.9819 & 0.9950\\
    Stacking  & 0.9644 & 0.0613 & 0.9633 & 0.9835 & 0.9840 & 1.0000\\
    BMC  & \textbf{0.9738} & 0.0437 & \textbf{0.9676} & \textbf{0.9942} &
    \textbf{0.9878} & 1.0000\\
  \end{tabular}
  \label{table:metrics_vvds}
\end{table*}

We apply the same four star-galaxy classification techniques
and four different combination methods,
and test them on the same test data.
We present in Table~\ref{table:metrics_vvds}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With poor training data, our data-driven methods,
\texttt{TPC} and \texttt{SOMc},
suffer a significant decrease in performance,
although they still in general perform better than 
the template fitting algorithm
\texttt{HB}.
The performance of our morphological separation method
does not depend on the training data,
and shows the best result in the MSE metric.
It also outperforms other base classifiers
in galaxy purity at galaxy completeness of 0.9964
and in star purity at star completeness of 0.7145.

Without sufficient training data,
the advantage of classification combination techniques
is more obvious.
Even WA, the simplest of combination techniques, outperforms
all individual classification techniques.
Interestingly, although TPC does not perform better
than morphological separation,
BoM still chooses TPC as the best model within each SOM cell
and shows exactly the same performance metrics as TPC.
Stacking and BMC techniques still outperform
all individual classification techniques,
and BMC's improvement in performance over TPC or mophological separation
is impressive.
Overall, the improvements are small but still significant
since these metrics are averaged over the full test data.

%% ROC curve for vvds i < 20
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/roc_cut.eps}
  \caption{A magnified view of the ROC curve of
           four star-galaxy individual classification approaches
           (TPC, SOMc, HB, and morphology)
           and the best-performing combination technique (BMC).
           trained with the reduced data set.
           We show only one of the four combination methods, BMC,
           which has the best overall performance.}
  \label{fig:roc_cut}
\end{figure}

In Figure~\ref{fig:roc_cut}, we plot the ROC curve
of all four individual classification algorithm
and our BMC technique for the poor training data case.
We note that a binary classifier,
such as our cut-based morphological separation,
whose output is either zero or one,
is a single point in the ROC space.
We vary the half-light radius to create a smooth ROC curve
for morphological separation.
It is clear that our BMC technique dominates
each individual classification techniques in the ROC space.

%% weights map for vvds i < 20
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/weights_cut.eps}
  \caption{A two-dimensional 10$\times$10 SOM representation
           showing the relative weights for the BMC combination technique
           applied to the four individual methods
           trained with the CFHTLenS data in the VVDS field
           with $i < 22$.}
  \label{fig:weights_cut}
\end{figure}

In Figure~\ref{fig:weights_cut}, we again show the $10\times10$
two-dimensional weight map defined by the SOM.
It seems that \texttt{TPC} performs well
in most of the parameter space.
When the quality of training data is relatively poor,
the performance of data driven algorithms will decrease,
while the performance of template fitting algorithms such as \text{HB}
is independent of training data.
Thus, it is perhaps not surprising that
BMC algorithm now uses information from \texttt{HB},
whereas it assigned almost no weight to \texttt{HB}
when we had a high-quality training data.
Another interesting pattern is that
\texttt{SOMc} and morphological separation seem complementary;
\texttt{SOMc} is weighted most strongly at fainter magnitudes,
while the morphological separation method
is weighted most strongly at brighter magnitudes.
It is not surprising that the morphological separation method
performs best at bright magnitudes,
and BMC uses almost no morphological information at faint magnitudes.

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
    \caption{Purity of morphological separation,
             TPC, and BMC techniques in 
             Table~\ref{table:metrics_vvds}
             as a function of the $i$-band magnitude.
             The top panel shows the fraction of stars
             as a function of magnitude.
             the galaxy and star purity for TPC and BMC techniques.
             The bottom two panels show
             the galaxy and star purity for morphological separation
             and BMC technique.
             We use the same threshold values from 
             Figure~\ref{fig:purity_mag} in the bottom four panels}.
    \label{fig:purity_mag}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_z.eps}
    \caption{The same figures as Figure~\ref{fig:purity_mag},
             but as a function of photometric redshift estimate
             (photo-$z$).}
    \label{fig:purity_z}
  \end{minipage}
\end{figure*}

We present the star and galaxy purity values
of Table~\ref{table:metrics_vvds} as a function of
$i$-band magnitude in Figure~\ref{fig:purity_mag}.
It is clear that BMC outperforms both
morphological separation and TPC classification techniques
over all magnitudes.
The galaxy purity of BMC is significantly better
than that of TPC at bright magnitudes, $i\lesssim21.5$,
while they are similar at faint magnitudes.
The star purity of BMC shows improvement over that of TPC
at faint magnitudes,
while they are similar at bright magnitudes.
As suggested by the weight maps in Figure~\ref{fig:weights_cut},
BMC can accomplish this by combining information from
all base classifiers,
\eg giving more weight to the morphological separation method
at bright magnitudes.
We can also see in the bottom two panels that
the performance of morphological separation suffers
at intermediate magnitudes $21\lesssim i \lesssim23$,
while BMC shows consistently better performance
over all magnitude ranges.

We also show the star and galaxy purity values
of Table~\ref{table:metrics_vvds} as a function of
photometric redshift in Figure~\ref{fig:purity_z}.
Photo-$z$ is estimated with
the \texttt{BPZ} algorithm~\citep{benitez2000bayesian}
and provided with
the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
It is again clear that BMC outperforms
both individual classification techniques over all redshift range.

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/p_dist_all.eps}
    \caption{Histogram of the posterior probability that
             a source is a star for morphological separation (top),
             TPC (middle), and BMC (bottom) techniques
             for the entire training data
             (the CFHTLenS data in the DEEP2, SDSS, VIPERS, and VVDS
             fields as discussed in Section~\ref{section:rich_training}).}
    \label{fig:p_dist_all}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/p_dist_cut.eps}
    \caption{The same figures as Figure~\ref{fig:p_dist_all},
             but for TPC and BMC trained with 
             poor training data
             (the CFHTLenS data in the VDDS field
             with a magnitude cut of $i<22$).}
    \label{fig:p_dist_cut}
  \end{minipage}
\end{figure*}

We show the distribution of posterior star probabilities
in Figures~\ref{fig:p_dist_all} and \ref{fig:p_dist_cut}.
The histogram of spectroscopic galaxies is in blue,
and the histogram of true stars is in red.
It is interesting that the BMC technique assigns
a probability $P\left(S\right)$ greater than 0.6
to almost no true galaxies in Figure~\ref{fig:p_dist_all},
presumably because BMC utilizes information from
different types of classification techniques.
This pattern is more pronounced in Figure~\ref{fig:p_dist_cut},
where the $P\left(S\right)$ distribution of BMC for true galaxies
falls off sharply at $P\left(S\right)\approx0.95$,
while morphological separation and TPC techniques
classify some true galaxies as stars with absolute certainty.

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figures/color_color.eps}
  \caption{Color-color diagrams of sources with
           $0.45<P\left(S\right)<0.55$ when the BMC technique
           is applied to the poor training data.}
  \label{fig:color_color}
\end{figure*}

To identify the regions in color-color space
where our BMC technique struggles to classify stars and galaxies,
we present the distribution over the color-color space for sources with
$0.45 < P(S) < 0.55$ in Figure~\ref{fig:color_color}.
Not surprisingly, these regions correspond to
where there is a significant overlap between stars and galaxies
in the color space.
It may be possible to improve the performance of
star-galaxy classification by obtaining more information
through spectroscopic follow-up, manual study, or citizen science projects,
or by developing a classification technique specifically designed
for these regions of the color space and
combining it with other techniques in a Bayesian framework.

\section{Conclusions}
  \label{section:conclusions}

When there's wealth of training data,
TPC is so good that combination techniques show
no significant improvement over TPC.

Even with poor (but reasonable) training data,
TPC outperforms unsupervised methods.

In a more realistic case where the training data is poor,
BMC should be used.

Probabilistic classification is better.
Binary classification can be transformed into a probabilistic
classification when combined with other models in a Bayesian framework.

\textcolor{red}{more here}

\section*{Acknowledgements}

We thank the referee for a careful reading of the manuscript
and comments that improved this work.
We thank Nacho Sevilla \textcolor{red}{Jacobo? anyone else?} for helpful and insightful conversations.
We acknowledge support from the National Science Grant No.
\textcolor{red}{XXXX}.

We gratefully acknowledge the use of computing resources
from the Extreme Science and Engineering Discovery Environment (XSEDE),
which is supported by National Science Foundation grant number
\textcolor{red}{XXXX}.

Funding for CFHTLenS has been provided by
\textcolor{red}{XXXX}.

\bibliographystyle{mn2e}

\bibliography{sg_paper}

\bsp

\label{lastpage}

\end{document}

### 2015-03-30 paper/main.tex
galaxy spectra are interpolated.

@@
 and G-K giant and supergiant stars.
+\textcolor{blue}{
+We supplement the stellar library with
+100 SEDs from \cite{chabrier2000evolutionary},
+which include low mass stars and brown dwarfs
+with different $T_{eff}$ and surface gravities.
+We also include four white dwarf templates of
+\cite{bohlin1995white}.
+}
 For our galaxy templates,
@@
 two star bursting galaxy templates from \cite{kinney1996template}.
-\textcolor{red}{interpolated}.
+\textcolor{blue}{
+From the six original CWW and Kinney spectra,
+intermediate templates are created by interpolation
+for a total of 51 SEDs.
+}
 

### 2015-03-30 paper/main.tex
remove table 1.

@@
 and its application to the estimation of
-photo-$z$ probability distribution funtions,
+photo-$z$ probability distribution functions,
 we refer the reader to \cite{carrascokind2014somz}.
@@
 based on its observed magnitudes is template fitting.
-Template fitting algorithms are unsupervised;
+Template fitting algorithms do not require a spectroscopic training sample;
 there is no need for additional knowledge outside the
@@
 by creating two-dimensional SOM representations of
-the full nine-dimensional magnitude-color space.
+the full nine-dimensional magnitude-color space:
+$u$, $g$, $r$, $i$, $z$, $u-g$, $g-r$, $r-i$, and $i-z$.
 A SOM representation can be rectangular, hexagonal, or spherical;
@@
 the fraction $\left(M_g + M_s\right) / N$,
-where $M_g$ are $M_s$ are defined in Table~\ref{table:def_numbers},
+where $M_g$ is the number of true galaxies classified as stars,
+$M_s$ is the number of true stars classified as galaxies,
 and $N$ is the total number of sources.
@@
 
-\begin{table}
-  \caption{The definition of the number of stars/galaxies
-           classified and misclassified as stars/galaxies.}
-  \centering
-  \begin{tabular}{ l | c | c }
-    \hline
-                           & True galaxies & True stars \\ \hline
-    Classified as galaxies & $N_g$ & $M_s$ \\ \hline
-    Classified as stars & $M_g$ & $N_s$ \\
-    \hline
-  \end{tabular}
-  \label{table:def_numbers}
-\end{table}
+%%\begin{table}
+%%  \caption{The definition of the number of stars/galaxies
+%%           classified and misclassified as stars/galaxies.}
+%%  \centering
+%%  \begin{tabular}{ l | c | c }
+%%    \hline
+%%                           & True galaxies & True stars \\ \hline
+%%    Classified as galaxies & $N_g$ & $M_s$ \\ \hline
+%%    Classified as stars & $M_g$ & $N_s$ \\
+%%    \hline
+%%  \end{tabular}
+%%  \label{table:def_numbers}
+%%\end{table}
 
@@
 \noindent
-where $N_g$ and $M_g$ are define in Table~\ref{table:def_numbers}.
+where $N_g$ is the number of true galaxies classified as galaxies
+and $M_g$ is the number of true galaxies classified as stars.
 We define the galaxy \textit{purity} $p_g$ (also known as precision
@@
 \noindent
-where $N_g$ and $M_g$ are define in Table~\ref{table:def_numbers}.
+where $M_s$ is the number of true stars classified as galaxies.
 Star completeness and purity are defined in a similar manner.

### 2015-03-31 paper/main.tex
add magnitude cut and z cut.

@@
 \usepackage{natbib}
+\usepackage{url}
 
@@
 
-%%\begin{table}
-%%  \caption{The definition of the number of stars/galaxies
-%%           classified and misclassified as stars/galaxies.}
-%%  \centering
-%%  \begin{tabular}{ l | c | c }
-%%    \hline
-%%                           & True galaxies & True stars \\ \hline
-%%    Classified as galaxies & $N_g$ & $M_s$ \\ \hline
-%%    Classified as stars & $M_g$ & $N_s$ \\
-%%    \hline
-%%  \end{tabular}
-%%  \label{table:def_numbers}
-%%\end{table}
-
 Although the space $\mathbf{E}$ of potential model combinations
@@
 with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
-It covers a total of 154 square degress
+It covers a total of 154 square degrees
 in the four fields of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
@@
   \centering
-%  \begin{tabular}{l c c c c c c}
-%    Classifier & AUC & MSE &
-%    $p_{g}\left(c_g=0.9964\right)$ & $p_{s}\left(c_s=0.7145\right)$ &
-%    $p_{g}\left(c_g=0.9600\right)$ & $p_{s}\left(c_s=0.2500\right)$ \\
-%    \hline
-%    TPC  & 0.9870 & 0.02081 & 0.9717 & 0.9843 & 0.9925 & 0.9977\\
-%    SOMc & 0.9683 & 0.04520 & 0.9099 & 0.8487 & 0.9781 & 0.9844\\
-%    HB   & 0.8954 & 0.08777 & 0.9048 & 0.7407 & 0.9593 & 0.7609\\
-%    Morphology   & -      & 0.0397 & 0.9597 & 0.9666 & - & -\\
-%    WA   & 0.9718 & 0.03292 & 0.9703 & 0.9876 & 0.9766 & 1.0000\\
-%    BoM  & 0.9870 & 0.02081 & 0.9733 & 0.9843 & 0.9925 & 0.9977\\
-%    Stacking  & \textbf{0.9878} & 0.02043 & 0.9729 & 0.9909 &
-%    \textbf{0.9928} & 1.0000\\
-%    BMC  & 0.9849 & \textbf{0.01911} & \textbf{0.9797} & \textbf{0.9950} &
-%    0.9916 & 1.0000\\
-%  \end{tabular}
   \input{tables/metrics_all.tex}
@@
 and often outperform \texttt{TPC}.
-We observe that our stacking algorithm shows the best performance
-as measured by the AUC,
-while BMC shows the best performance
-as measured by the MSE.
+\textcolor{blue}{
+Although \texttt{TPC} shows the best performance as measured by the AUC,
+BMC shows the best performance in all other metrics.
+}
 
@@
     \caption{A two-dimensional 10$\times$10 SOM representation
-             showing the mean $i$-band magnitude (top left) and
-             the mean values of three colors,
-             $u-g$, $g-r$, and $i-z$,
+             showing the mean $i$-band magnitude (top left),
+             the fraction of true stars in each cell (top right),
+             and the mean values of $u-g$ and $g-r$
              for the cross validation data.}
@@
 In Figure~\ref{fig:som_colors}, we show
-the mean CFHTLenS $i$-band magnitude in the top right panel,
-and the mean $u-g$, $g-r$, and $i-z$ colors in the remaining three panels.
+the mean CFHTLenS $i$-band magnitude in the top left panel
+\textcolor{blue}{
+and the fraction of stars in each cell in the top right panel.
+}
+The bottom two panels show the mean $u-g$ and  $g-r$ colors.
 These two-dimensional maps clearly shows
-the ability of the SOM to preserve
-relationships between sources
-when it projects the full nine-dimensional space to
-the two-dimensional map.
-The SOM mapping is a non-linear representation of
-all magnitudes and colors,
-and thus the CHFTLenS $i$-band magnitude and color maps should 
-only be used to provide guidance.
+the ability of the SOM to preserve relationships between sources
+when it projects the full nine-dimensional space to the two-dimensional map.
+The SOM mapping is a non-linear representation of all magnitudes and colors,
+and thus these SOM maps should only be used to provide guidance.
 
@@
 Not surprisingly, the morphological separation
-performs best in the lower right corner of the $i$-band magnitude map,
+performs best in the top left corner of the $i$-band magnitude map,
 which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
@@
 with $0<u-g<0.5$ and $0<g-r<0.5$.
-The BMC technique assigns almost no weight to
-the worst-performing algorithm, \texttt{HB},
-when there is a wealth of training data.
-
-\subsection{Insufficient Training Data}
-  \label{section:poor_training}
+\textcolor{blue}{
+Although \texttt{HB} shows the worst performance
+when there exists ample training data,
+BMC still utilizes information from \texttt{HB}
+to produce an overall improvement in performance,
+especially at intermediate magnitudes $20\gtrsim i \lesssim20$.
+}
+
+\subsection{The Quality of Training Data}
+  \label{section:quality_training}
 
@@
 Thus, we investigate the impact of training set quality
-by considering a more realistic case
+by gradually decreasing the size of the training set.
+
+\begin{figure*}
+  \begin{minipage}[b]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
+    \caption{The classification performance metrics for TPC, HB, morphology,
+      and BMC as applied to the CFHTLenS data in the VVDS field
+      with various magnitude cuts.
+      We show only one of the four combination methods, BMC,
+      which has the best overall performance.}
+    \label{fig:perform_mag_cut}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
+    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
+      $z_{\text{spec}}$ cuts.}
+    \label{fig:perform_z_cut}
+  \end{minipage}
+\end{figure*}
+
+\subsection{Insufficient Training Data}
+  \label{section:poor_training}
+
+We also investigate the impact of training set quality
+by analyzing in detail a more realistic case
 where the training data is available
@@
 To emulate this scenario,
-we only use objects that have spectroscopic labels from the VVDS
+we only use objects that have spectroscopic labels
+from the VVDS 0226-04 field (which is located within the CFHTLS W1 field)
 and impose a magnitude cut of $i < 22.0$ in the training data,
@@
 
+\textcolor{blue}{begin new text}
+
+We have presented and analyzed a novel star-galaxy classification framework
+for combining classifications from the CFHTLenS project.
+In particular, we use four independent star-galaxy classification techniques:
+a morphological separation method;
+\texttt{TPC}, a supervised machine learning technique
+based on prediction trees and a random forest;
+\texttt{SOMc}, an unsupervised machine leraning approach
+based on self-organizing maps and a random atlas;
+and \texttt{HB}, a hierarichical bayesian template-fitting method
+that we have modified and parallized.
+Both \texttt{TPC} and \texttt{SOMc} are currently availble within
+a software package named
+\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}.
+Our implementaion of \texttt{HB} and \texttt{BMC},
+as well as IPython notebooks that were used to
+produce the results in this paper,
+are available at \url{https://github.com/EdwardJKim/astroclass}.
+
+\textcolor{blue}{end new text}
+
 When there's wealth of training data,
@@
 
+We note that \cite{carrascokind2014exhausting} analyzed
+different techniques for combining
+photometric redshift probability density functions (photo-$z$ PDFs),
+and also found that BMC is in general the best
+photo-$z$ PDF combination technique.
+
 \textcolor{red}{more here}
@@
 and comments that improved this work.
-We thank Nacho Sevilla \textcolor{red}{Jacobo? anyone else?} for helpful and insightful conversations.
+We thank Ignacio Sevilla for helpful and insightful conversations.
 We acknowledge support from the National Science Grant No.

### 2015-04-13 paper/main.tex
add purity vs mag plot.

@@
 \begin{figure*}
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[t]{0.49\linewidth}
     \includegraphics[width=\textwidth]{figures/som_colors.eps}
@@
   \hfill
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[t]{0.49\linewidth}
     \includegraphics[width=\textwidth]{figures/weights.eps}
@@
 to produce an overall improvement in performance,
-especially at intermediate magnitudes $20\gtrsim i \lesssim20$.
+especially at intermediate magnitudes $20\gtrsim i \lesssim22$.
 }
+
+In Figure~\ref{fig:purity_mag}, 
+we compare the star and galaxy purity values
+for morphological separation, TPC, and BMC techniques
+as a function of $i$-band magnitude.
+Although morphological separation shows a slightly better performance
+at bright magnitudes $i < 21$,
+BMC outperforms both morphological separation and TPC
+at faint magnitudes $i > 21$.
+As shown in the histogram, the histogram peaks at $i \sim 22$,
+and BMC therefore outperforms the morphological separation technique
+for the majority of objects.
+It is also clear that BMC outperforms TPC over all magnitudes.
+BMC can accomplish this by combining information from
+all base classifiers,
+\eg giving more weight to the morphological separation method
+at bright magnitudes.
+
+\begin{figure}
+%%  \begin{minipage}[t]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
+    \caption{Purity of morphological separation,
+             TPC, and BMC techniques in 
+             Table~\ref{table:metrics_vvds}
+             as a function of the $i$-band magnitude.
+             The top panel shows the histogram of objects in the test data set.
+             The second panel shows the fraction of stars
+             as a function of magnitude. The error bars are Poisson.
+             The bottom two panels compares
+             the galaxy and star purity for morphological separation, TPC,
+             and BMC techniques as a function of magnitude.
+             The error bars for completeness and purity are not shown for
+             clarity, but they are approximately 10 percent.
+             }.
+    \label{fig:purity_mag}
+\end{figure}
 
@@
 Thus, we investigate the impact of training set quality
-by gradually decreasing the size of the training set.
+by imposing various magnitude and $z$ cuts
+to decrease the size of the training set.
+
+In Figure.~\ref{fig:perform_mag_cut}, we present
+a visual comparison between different classification techniques
+when various magnitude cuts are applied.
+
+In Figure.~\ref{fig:perform_z_cut}, we also present
+the result of imposing various $z$ cuts.
 
 \begin{figure*}
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[t]{0.49\linewidth}
     \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
@@
       with various magnitude cuts.
+      The top panel shows the number of sources in the training set
+      at corresponding magnitude cuts.
       We show only one of the four combination methods, BMC,
@@
   \hfill
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[t]{0.49\linewidth}
     \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
@@
 and impose a magnitude cut of $i < 22.0$ in the training data,
-leaving us a training set with 6,361 objects.
+leaving us a training set with 1,365 objects.
 
@@
   \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
+    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
     \caption{Purity of morphological separation,
@@
              Figure~\ref{fig:purity_mag} in the bottom four panels}.
-    \label{fig:purity_mag}
+    \label{fig:purity_mag_cut}
   \end{minipage}

### 2015-04-15 paper/main.tex
fix purity vs mag figure; add purity vs z.
update up to figure 8.

@@
 \usepackage{url}
+\usepackage{float}
 
@@
 \begin{abstract}
-\textcolor{red}{Abstract goes here.}
+\textcolor{blue}{
+Star-galaxy separation in currently ongoing and future
+large photometric surveys, such as DES and LSST,
+will detect a vast number of unresolved galaxies at faint magnitudes.
+There exist a variety of star-galaxy classification techniques,
+each with their own strenths and weaknesses.
+In this paper, we present a novel hybrid, ensemble classification
+framework that combines and fully exploits different techniques
+to produce a more robust classification.
+To demonstrate this hybrid technique,
+we combine a purely morphological classifier,
+a supervised machine learning technique based on random forest,
+an unsupervised method based on self-organizing maps,
+and a heirarchical Bayeisan template fitting algorithm.
+Any of these four methods could
+easily be replaced by a similar method
+or be extended to include other existing techniques.
+Testing with CFHTLenS data,
+we demonstrate that our combination technique improves
+the overall performance over any individual classification method.
+}
+
 \end{abstract}
@@
 
-In Figure~\ref{fig:purity_mag}, 
+%\begin{figure*}
+%  \begin{minipage}[t]{0.49\linewidth}
+%    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
+%    \caption{Purity of morphological separation,
+%             TPC, and BMC techniques in 
+%             as a function of the $i$-band magnitude.
+%             The top panel shows the histogram of objects in the test data set.
+%             The second panel shows the fraction of stars
+%             as a function of magnitude. The error bars are Poisson.
+%             The bottom two panels compare
+%             the galaxy and star purity for morphological separation, TPC,
+%             and BMC techniques as a function of magnitude.
+%             The error bars for completeness and purity are not shown for
+%             clarity, but they are approximately 10 percent.}
+%    \label{fig:purity_mag}
+%  \end{minipage}
+%  \hfill
+%  \begin{minipage}[t]{0.49\linewidth}
+%    \includegraphics[width=\textwidth]{figures/purity_z.eps}
+%    \caption{The same figures as Figure~\ref{fig:purity_mag},
+%             but as a function of photometric redshift estimate
+%             (photo-$z$).}
+%    \label{fig:purity_z}
+%  \end{minipage}
+%\end{figure*}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_mag.eps}
+  \caption{Purity of morphological separation,
+           TPC, and BMC techniques in 
+           as a function of the $i$-band magnitude.
+           The top panel shows the histogram of objects in the test data set.
+           The second panel shows the fraction of stars
+           as a function of magnitude. The error bars are Poisson.
+           The bottom two panels compare
+           the galaxy and star purity for morphological separation, TPC,
+           and BMC techniques as a function of magnitude.
+           The error bars for completeness and purity are not shown for
+           clarity, but they are approximately 10 percent.}
+  \label{fig:purity_mag}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_z.eps}
+  \caption{The same figures as Figure~\ref{fig:purity_mag},
+           but as a function of photometric redshift estimate
+           (photo-$z$).}
+  \label{fig:purity_z}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_g-r.eps}
+  \caption{The same figure as Figure~\ref{purity_mag},
+           but as a function of $g-r$ color.}
+  \label{fig:purity_g-r}
+\end{figure}
+
+\textcolor{blue}{In Figure~\ref{fig:purity_mag}, 
 we compare the star and galaxy purity values
@@
 at faint magnitudes $i > 21$.
-As shown in the histogram, the histogram peaks at $i \sim 22$,
+As shown in Figure.~\ref{fig:purity_mag},
+the histogram of distribution of objects peaks at $i \sim 22$,
 and BMC therefore outperforms the morphological separation technique
@@
 It is also clear that BMC outperforms TPC over all magnitudes.
-BMC can accomplish this by combining information from
+BMC can presumably accomplish this by combining information from
 all base classifiers,
 \eg giving more weight to the morphological separation method
-at bright magnitudes.
-
-\begin{figure}
-%%  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
-    \caption{Purity of morphological separation,
-             TPC, and BMC techniques in 
-             Table~\ref{table:metrics_vvds}
-             as a function of the $i$-band magnitude.
-             The top panel shows the histogram of objects in the test data set.
-             The second panel shows the fraction of stars
-             as a function of magnitude. The error bars are Poisson.
-             The bottom two panels compares
-             the galaxy and star purity for morphological separation, TPC,
-             and BMC techniques as a function of magnitude.
-             The error bars for completeness and purity are not shown for
-             clarity, but they are approximately 10 percent.
-             }.
-    \label{fig:purity_mag}
-\end{figure}
-
-\subsection{The Quality of Training Data}
-  \label{section:quality_training}
+at bright magnitudes.}
+
+\textcolor{blue}{We also show the star and galaxy purity values
+of Table~\ref{table:metrics_all} as a function of
+photometric redshift in Figure~\ref{fig:purity_z}.
+Photo-$z$ is estimated with
+the \texttt{BPZ} algorithm~\citep{benitez2000bayesian}
+and provided with
+the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
+Although the morphological separation method outperformed BMC
+at bright magnitudes in Figure.~\ref{fig:purity_mag},
+it is clear that BMC outperforms
+both morphological separation and TPC over all redshift range.}
+
+\textcolor{blue}{In Figure.~\ref{fig:purity_g-r},
+we present how the star and galaxy purity values vary
+as a function of $g-r$ color.
+It is again clear that BMC outperforms
+both morphological separation and TPC over all $g-r$ color range.  }
+
+\subsection{Heterogeneous Training}
+  \label{section:poor_training}
 
@@
 Thus, we investigate the impact of training set quality
-by imposing various magnitude and $z$ cuts
-to decrease the size of the training set.
-
-In Figure.~\ref{fig:perform_mag_cut}, we present
-a visual comparison between different classification techniques
-when various magnitude cuts are applied.
-
-In Figure.~\ref{fig:perform_z_cut}, we also present
-the result of imposing various $z$ cuts.
-
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
-    \caption{The classification performance metrics for TPC, HB, morphology,
-      and BMC as applied to the CFHTLenS data in the VVDS field
-      with various magnitude cuts.
-      The top panel shows the number of sources in the training set
-      at corresponding magnitude cuts.
-      We show only one of the four combination methods, BMC,
-      which has the best overall performance.}
-    \label{fig:perform_mag_cut}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
-    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
-      $z_{\text{spec}}$ cuts.}
-    \label{fig:perform_z_cut}
-  \end{minipage}
-\end{figure*}
-
-\subsection{Insufficient Training Data}
-  \label{section:poor_training}
-
-We also investigate the impact of training set quality
 by analyzing in detail a more realistic case
@@
 leaving us a training set with 1,365 objects.
+We apply the same four star-galaxy classification techniques
+and four different combination methods,
+and test them on the same test data.
+Since the demographics of sources in the test data set
+does not show the same distribution as the training data set,
+this also serves as a test of the efficacy of heterogenous training data.
 
@@
            and the four different classification combination methods
-           as applied to the CFHTLenS data in the VVDS field
+           as applied to the CFHTLenS data in the VVDS W1 field
            with a magnitude cut of $i < 22.0$.
@@
   \centering
-  \begin{tabular}{l c c c c c c}
-    Classifier & AUC & MSE &
-    $p_{g}\left(c_g=0.9964\right)$ & $p_{s}\left(c_s=0.7145\right)$ &
-    $p_{g}\left(c_g=0.9600\right)$ & $p_{s}\left(c_s=0.2500\right)$\\
-    \hline
-    TPC  & 0.9601 & 0.0728 & 0.9438 & 0.9329 & 0.9819 & 0.9950\\
-    SOMc & 0.9008 & 0.1679 & 0.8843 & 0.5044 & 0.9244 & 0.7726\\
-    HB   & 0.8960 & 0.0894 & 0.9074 & 0.7516 & 0.9615 & 0.7793\\
-    Morphology   & -      & \textbf{0.0397} & 0.9597 & 0.9666 & - & -\\
-    WA   & 0.9639 & 0.0581 & 0.9613 & 0.9787 & 0.9756 & 1.0000\\
-    BoM  & 0.9601 & 0.0728 & 0.9438 & 0.9329 & 0.9819 & 0.9950\\
-    Stacking  & 0.9644 & 0.0613 & 0.9633 & 0.9835 & 0.9840 & 1.0000\\
-    BMC  & \textbf{0.9738} & 0.0437 & \textbf{0.9676} & \textbf{0.9942} &
-    \textbf{0.9878} & 1.0000\\
-  \end{tabular}
-  \label{table:metrics_vvds}
+  \input{tables/metrics_cut.tex}
+  \label{table:metrics_cut}
 \end{table*}
 
-We apply the same four star-galaxy classification techniques
-and four different combination methods,
-and test them on the same test data.
-We present in Table~\ref{table:metrics_vvds}
+We present in Table~\ref{table:metrics_cut}
 the same six metrics for each method,
@@
 %% ROC curve for vvds i < 20
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/roc_cut.eps}
-  \caption{A magnified view of the ROC curve of
-           four star-galaxy individual classification approaches
-           (TPC, SOMc, HB, and morphology)
-           and the best-performing combination technique (BMC).
-           trained with the reduced data set.
-           We show only one of the four combination methods, BMC,
-           which has the best overall performance.}
-  \label{fig:roc_cut}
-\end{figure}
-
-In Figure~\ref{fig:roc_cut}, we plot the ROC curve
-of all four individual classification algorithm
-and our BMC technique for the poor training data case.
-We note that a binary classifier,
-such as our cut-based morphological separation,
-whose output is either zero or one,
-is a single point in the ROC space.
-We vary the half-light radius to create a smooth ROC curve
-for morphological separation.
-It is clear that our BMC technique dominates
-each individual classification techniques in the ROC space.
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=\linewidth]{figures/roc_cut.eps}
+%  \caption{A magnified view of the ROC curve of
+%           four star-galaxy individual classification approaches
+%           (TPC, SOMc, HB, and morphology)
+%           and the best-performing combination technique (BMC).
+%           trained with the reduced data set.
+%           We show only one of the four combination methods, BMC,
+%           which has the best overall performance.}
+%  \label{fig:roc_cut}
+%\end{figure}
+
+%In Figure~\ref{fig:roc_cut}, we plot the ROC curve
+%of all four individual classification algorithm
+%and our BMC technique for the poor training data case.
+%We note that a binary classifier,
+%such as our cut-based morphological separation,
+%whose output is either zero or one,
+%is a single point in the ROC space.
+%We vary the half-light radius to create a smooth ROC curve
+%for morphological separation.
+%It is clear that our BMC technique dominates
+%each individual classification techniques in the ROC space.
 
 %% weights map for vvds i < 20
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/weights_cut.eps}
-  \caption{A two-dimensional 10$\times$10 SOM representation
-           showing the relative weights for the BMC combination technique
-           applied to the four individual methods
-           trained with the CFHTLenS data in the VVDS field
+
+\begin{figure*}
+  \begin{minipage}[t]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/som_colors_cut.eps}
+    \caption{The same figure as Figure~\ref{fig:som_colors_cut}
+        but for the BMC technique applied to the four individual methods
+        trained with the CFHTLenS data in the VVDS W1 field
+        with $i < 22$.}
+    \label{fig:som_colors_cut}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/weights_cut.eps}
+    \caption{The same figure as FIgure~\ref{fig:weights_cut}
+        but for the BMC technique applied to the four invidiual methods
+        trained with the CFHTLenS data in the VVDS W1 field
         with $i < 22$.}
     \label{fig:weights_cut}
-\end{figure}
+  \end{minipage}
+\end{figure*}
 
@@
 
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
-    \caption{Purity of morphological separation,
-             TPC, and BMC techniques in 
-             Table~\ref{table:metrics_vvds}
-             as a function of the $i$-band magnitude.
-             The top panel shows the fraction of stars
-             as a function of magnitude.
-             the galaxy and star purity for TPC and BMC techniques.
-             The bottom two panels show
-             the galaxy and star purity for morphological separation
-             and BMC technique.
-             We use the same threshold values from 
-             Figure~\ref{fig:purity_mag} in the bottom four panels}.
-    \label{fig:purity_mag_cut}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/purity_z.eps}
-    \caption{The same figures as Figure~\ref{fig:purity_mag},
-             but as a function of photometric redshift estimate
-             (photo-$z$).}
-    \label{fig:purity_z}
-  \end{minipage}
-\end{figure*}
+%\begin{figure*}
+%  \begin{minipage}[t]{0.49\linewidth}
+%    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
+%    \caption{Purity of morphological separation,
+%             TPC, and BMC techniques in 
+%             Table~\ref{table:metrics_vvds}
+%             as a function of the $i$-band magnitude.
+%             The top panel shows the fraction of stars
+%             as a function of magnitude.
+%             the galaxy and star purity for TPC and BMC techniques.
+%             The bottom two panels show
+%             the galaxy and star purity for morphological separation
+%             and BMC technique.
+%             We use the same threshold values from 
+%             Figure~\ref{fig:purity_mag} in the bottom four panels}.
+%    \label{fig:purity_mag_cut}
+%  \end{minipage}
+%  \hfill
+%  \begin{minipage}[t]{0.49\linewidth}
+%    \includegraphics[width=\textwidth]{figures/purity_z.eps}
+%    \caption{The same figures as Figure~\ref{fig:purity_mag},
+%             but as a function of photometric redshift estimate
+%             (photo-$z$).}
+%    \label{fig:purity_z}
+%  \end{minipage}
+%\end{figure*}
 
@@
 
+\subsection{The Quality of Training Data}
+  \label{section:quality_training}
+
+It is very costly in terms of telescope time to
+obtain a large sample spectroscopic observations
+down to the limiting magnitude of a photometric sample.
+Thus, we investigate the impact of training set quality
+by imposing various magnitude and $z$ cuts
+to decrease the size of the training set.
+
+In Figure.~\ref{fig:perform_mag_cut}, we present
+a visual comparison between different classification techniques
+when various magnitude cuts are applied.
+
+In Figure.~\ref{fig:perform_z_cut}, we also present
+the result of imposing various $z$ cuts.
+
+\begin{figure*}
+  \begin{minipage}[t]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
+    \caption{The classification performance metrics for TPC, HB, morphology,
+      and BMC as applied to the CFHTLenS data in the VVDS field
+      with various magnitude cuts.
+      The top panel shows the number of sources in the training set
+      at corresponding magnitude cuts.
+      We show only one of the four combination methods, BMC,
+      which has the best overall performance.}
+    \label{fig:perform_mag_cut}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{0.49\linewidth}
+    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
+    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
+      $z_{\text{spec}}$ cuts.}
+    \label{fig:perform_z_cut}
+  \end{minipage}
+\end{figure*}
 \section{Conclusions}
@@
 
-\textcolor{blue}{begin new text}
-
-We have presented and analyzed a novel star-galaxy classification framework
+\textcolor{blue}{We
+have presented and analyzed a novel star-galaxy classification framework
 for combining classifications from the CFHTLenS project.
@@
 produce the results in this paper,
-are available at \url{https://github.com/EdwardJKim/astroclass}.
-
-\textcolor{blue}{end new text}
+are available at \url{https://github.com/EdwardJKim/astroclass}.  }
 

### 2015-04-15 paper/main.tex
Add purity in vvds w1 figures.

@@
 
-%\begin{figure*}
-%  \begin{minipage}[t]{0.49\linewidth}
-%    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
-%    \caption{Purity of morphological separation,
-%             TPC, and BMC techniques in 
-%             as a function of the $i$-band magnitude.
-%             The top panel shows the histogram of objects in the test data set.
-%             The second panel shows the fraction of stars
-%             as a function of magnitude. The error bars are Poisson.
-%             The bottom two panels compare
-%             the galaxy and star purity for morphological separation, TPC,
-%             and BMC techniques as a function of magnitude.
-%             The error bars for completeness and purity are not shown for
-%             clarity, but they are approximately 10 percent.}
-%    \label{fig:purity_mag}
-%  \end{minipage}
-%  \hfill
-%  \begin{minipage}[t]{0.49\linewidth}
-%    \includegraphics[width=\textwidth]{figures/purity_z.eps}
-%    \caption{The same figures as Figure~\ref{fig:purity_mag},
-%             but as a function of photometric redshift estimate
-%             (photo-$z$).}
-%    \label{fig:purity_z}
-%  \end{minipage}
-%\end{figure*}
-
 \begin{figure}
@@
   \caption{Purity of morphological separation,
-           TPC, and BMC techniques in 
+           TPC, and BMC techniques
            as a function of the $i$-band magnitude.
@@
     \includegraphics[width=\textwidth]{figures/som_colors_cut.eps}
-    \caption{The same figure as Figure~\ref{fig:som_colors_cut}
+    \caption{The same figure as Figure~\ref{fig:som_colors}
         but for the BMC technique applied to the four individual methods
@@
     \includegraphics[width=\textwidth]{figures/weights_cut.eps}
-    \caption{The same figure as FIgure~\ref{fig:weights_cut}
+    \caption{The same figure as Figure~\ref{fig:weights}
         but for the BMC technique applied to the four invidiual methods
@@
 two-dimensional weight map defined by the SOM.
-It seems that \texttt{TPC} performs well
-in most of the parameter space.
-When the quality of training data is relatively poor,
+%It seems that \texttt{TPC} performs well
+%in most of the parameter space.
+\textcolor{blue}{When the quality of training data is relatively poor,
 the performance of data driven algorithms will decrease,
-while the performance of template fitting algorithms such as \text{HB}
+while the performance of template fitting algorithms such as \texttt{HB}
 is independent of training data.
-Thus, it is perhaps not surprising that
-BMC algorithm now uses information from \texttt{HB},
-whereas it assigned almost no weight to \texttt{HB}
-when we had a high-quality training data.
-Another interesting pattern is that
-\texttt{SOMc} and morphological separation seem complementary;
-\texttt{SOMc} is weighted most strongly at fainter magnitudes,
-while the morphological separation method
-is weighted most strongly at brighter magnitudes.
+Thus, comparing the weight maps of Figure~\ref{fig:weights}
+and Figure~\ref{fig:weights_cut},
+it is perhaps not surprising that
+BMC algorithm now uses more information from
+morphological separation and \texttt{HB},
+while it uses considerably less information from TPC.
+%Another interesting pattern is that
+%\texttt{SOMc} and morphological separation seem complementary;
+%\texttt{SOMc} is weighted most strongly at fainter magnitudes,
+%while the morphological separation method
+%is weighted most strongly at brighter magnitudes.
 It is not surprising that the morphological separation method
 performs best at bright magnitudes,
-and BMC uses almost no morphological information at faint magnitudes.
-
-%\begin{figure*}
-%  \begin{minipage}[t]{0.49\linewidth}
-%    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
-%    \caption{Purity of morphological separation,
-%             TPC, and BMC techniques in 
-%             Table~\ref{table:metrics_vvds}
-%             as a function of the $i$-band magnitude.
-%             The top panel shows the fraction of stars
-%             as a function of magnitude.
-%             the galaxy and star purity for TPC and BMC techniques.
-%             The bottom two panels show
-%             the galaxy and star purity for morphological separation
-%             and BMC technique.
-%             We use the same threshold values from 
-%             Figure~\ref{fig:purity_mag} in the bottom four panels}.
-%    \label{fig:purity_mag_cut}
-%  \end{minipage}
-%  \hfill
-%  \begin{minipage}[t]{0.49\linewidth}
-%    \includegraphics[width=\textwidth]{figures/purity_z.eps}
-%    \caption{The same figures as Figure~\ref{fig:purity_mag},
-%             but as a function of photometric redshift estimate
-%             (photo-$z$).}
-%    \label{fig:purity_z}
-%  \end{minipage}
-%\end{figure*}
+and BMC assigns more weight to \texttt{HB} at fainter magnitudes.}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_mag_cut.eps}
+  \caption{Purity as a function of the $i$-band magnitude
+           for morphological separation, TPC, and BMC techniques
+           in the VVDS W1 field with $i < 22$.
+           We repat in the top panel shows the histogram of objects
+           in the test data set.
+           The second panel shows the fraction of stars
+           in the VVDS W1 and W4 fields.
+           The error bars are Poisson.
+           The bottom two panels compare
+           the galaxy and star purity for morphological separation, TPC,
+           and BMC techniques as a function of magnitude.
+           The error bars for completeness and purity are not shown for
+           clarity, but they are approximately 10 percent.}
+  \label{fig:purity_mag_cut}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_z_cut.eps}
+  \caption{The same figures as Figure~\ref{fig:purity_mag_cut},
+           but as a function of photo-$z$.}
+  \label{fig:purity_z_cut}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_g-r_cut.eps}
+  \caption{The same figure as Figure~\ref{purity_mag_cut},
+           but as a function of $g-r$ color.}
+  \label{fig:purity_g-r_cut}
+\end{figure}
 
@@
 of Table~\ref{table:metrics_vvds} as a function of
-$i$-band magnitude in Figure~\ref{fig:purity_mag}.
+$i$-band magnitude in Figure~\ref{fig:purity_mag_cut}.
 It is clear that BMC outperforms both

### 2015-04-23 paper/main.tex
update draft.

@@
 \usepackage{url}
-\usepackage{float}
+\usepackage{placeins}
 
@@
 we demonstrate that our combination technique improves
-the overall performance over any individual classification method.
-}
+the overall performance over any individual classification method.}
+
+\textcolor{red}{more abstract.}
 
@@
 
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/som_colors.eps}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/som_colors.eps}
   \caption{A two-dimensional 10$\times$10 SOM representation
@@
   \label{fig:som_colors}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/weights.eps}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/weights.eps}
   \caption{A two-dimensional 10$\times$10 SOM representation
@@
   \label{fig:weights}
-  \end{minipage}
-\end{figure*}
+\end{figure}
 
@@
 especially at intermediate magnitudes $20\gtrsim i \lesssim22$.
-}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/purity_mag.eps}
-  \caption{Purity of morphological separation,
-           TPC, and BMC techniques
-           as a function of the $i$-band magnitude.
-           The top panel shows the histogram of objects in the test data set.
-           The second panel shows the fraction of stars
-           as a function of magnitude. The error bars are Poisson.
-           The bottom two panels compare
-           the galaxy and star purity for morphological separation, TPC,
-           and BMC techniques as a function of magnitude.
-           The error bars for completeness and purity are not shown for
-           clarity, but they are approximately 10 percent.}
-  \label{fig:purity_mag}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/purity_z.eps}
-  \caption{The same figures as Figure~\ref{fig:purity_mag},
-           but as a function of photometric redshift estimate
-           (photo-$z$).}
-  \label{fig:purity_z}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/purity_g-r.eps}
-  \caption{The same figure as Figure~\ref{purity_mag},
-           but as a function of $g-r$ color.}
-  \label{fig:purity_g-r}
-\end{figure}
+Another interesting pattern is that
+the four techniques seem complementary,
+and they are weighted most strongly in different regions
+of the SOM representations.}
 
@@
 for morphological separation, TPC, and BMC techniques
-as a function of $i$-band magnitude.
+as a function of $i$-band magnitude for the differential counts.
+We use the kernel density estimation \cite[KDE;][]{silverman1986density}.
 Although morphological separation shows a slightly better performance
@@
 at faint magnitudes $i > 21$.
-As shown in Figure.~\ref{fig:purity_mag},
-the histogram of distribution of objects peaks at $i \sim 22$,
+As the top panel shows,
+the number count distribution peaks at $i \sim 22$,
 and BMC therefore outperforms the morphological separation technique
@@
 \eg giving more weight to the morphological separation method
-at bright magnitudes.}
+at bright magnitudes.
+In Figure~\ref{fig:purity_mag_itnegrated}, we show
+the overall galaxy and star purity as a function of magnitude
+for the integrated counts.
+BMC is able to maintain a galaxy purity of 0.980
+up to $i \sim 24.5$, while the galaxy purity of
+the morphological separation method drops to 95.7\%
+beyond $i \lesssim 21$.}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_mag.pdf}
+  \caption{Purity as a function of the $i$-band magnitude
+           as estimated by the kernel density estimation (KDE) method
+           for the differential counts.
+           The top panel shows the histogram with a bin size of 0.1 mag
+           and the KDE for objects in the test set.
+           The second panel shows the fraction of stars estimated by KDE
+           as a function of magnitude.
+           The bottom two panels compare
+           the galaxy and star purity for BMC, TPC, and 
+           morphological separation as a function of magnitude.
+           Results for BMC, TPC, and morphological separation are in
+           red, green, and blue, respectively.
+           The $1 \sigma$ confidence bands are estaimted by
+           bootstrap resampling.}
+  \label{fig:purity_mag}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_mag_integrated.pdf}
+  \caption{Purity as a function of the $i$-band magnitude
+           for the integrated counts.
+           The bottom two panels compare
+           the galaxy and star purity for BMC, TPC, and 
+           morphological separation as a function of magnitude.
+           Results for BMC, TPC, and morphological separation are in
+           red, green, and blue, respectively.
+           The $1 \sigma$ error bars are Poisson.}
+  \label{fig:purity_mag_integrated}
+\end{figure}
 
@@
 the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
+The advantage of BMC over morphological separation and TPC is
+now more pronounced in Figure~\ref{fig:purity_z}.
 Although the morphological separation method outperformed BMC
@@
 it is clear that BMC outperforms
-both morphological separation and TPC over all redshift range.}
-
-\textcolor{blue}{In Figure.~\ref{fig:purity_g-r},
-we present how the star and galaxy purity values vary
+both morphological separation and TPC over all redshift range.
+We also present in Figure.~\ref{fig:purity_g-r}
+how the star and galaxy purity values vary
 as a function of $g-r$ color.
@@
 both morphological separation and TPC over all $g-r$ color range.  }
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_z.pdf}
+  \caption{The same figures as Figure~\ref{fig:purity_mag},
+           but as a function of photometric redshift estimate
+           (photo-$z$).
+           The bin size of histogram in the top panel is 0.02.}
+  \label{fig:purity_z}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_g-r.pdf}
+  \caption{The same figure as Figure~\ref{purity_mag},
+           but as a function of $g-r$ color.
+           The bin size of histogram in the top panel is 0.05.}
+  \label{fig:purity_g-r}
+\end{figure}
+
+\textcolor{blue}{We show the distribution of posterior star probabilities
+in Figures~\ref{fig:p_dist}.
+The histogram of spectroscopic galaxies is in red,
+and the histogram of true stars is in black.
+It is interesting that the BMC technique assigns
+a probability $P\left(S\right)$ greater than 0.75
+to significally fewer true galaxies than TPC,
+presumably because BMC utilizes information from
+different types of classification techniques.}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/p_dist.pdf}
+  \caption{Histogram of the posterior probability that
+           a source is a star for morphological separation (top),
+           TPC (middle), and BMC (bottom) techniques
+           for a high-quality training data.
+           The bin size is 0.05.}
+  \label{fig:p_dist_all}
+\end{figure}
 
@@
 Thus, we investigate the impact of training set quality
-by analyzing in detail a more realistic case
+by considering a more realistic case
 where the training data is available
-only for a small number of objects at bright magnitudes.
-To emulate this scenario,
+only for a small number of objects with bright magnitudes.
+\textcolor{blue}{ To emulate this scenario,
 we only use objects that have spectroscopic labels
@@
 Since the demographics of sources in the test data set
-does not show the same distribution as the training data set,
-this also serves as a test of the efficacy of heterogenous training data.
+are different from the distribution of the training data set,
+this also serves as a test of the efficacy of heterogenous training data.}
 
@@
 since these metrics are averaged over the full test data.
+\textcolor{red}{Note: Numbers for WA, BOM, and Stacking seem a little bit off.
+Will investigate.}
 
@@
 
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/som_colors_cut.eps}
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/som_colors_cut.eps}
   \caption{The same figure as Figure~\ref{fig:som_colors}
@@
   \label{fig:som_colors_cut}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/weights_cut.eps}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/weights_cut.eps}
   \caption{The same figure as Figure~\ref{fig:weights}
@@
   \label{fig:weights_cut}
-  \end{minipage}
-\end{figure*}
+\end{figure}
 
@@
 is independent of training data.
-Thus, comparing the weight maps of Figure~\ref{fig:weights}
-and Figure~\ref{fig:weights_cut},
-it is perhaps not surprising that
+Thus, when the weight maps of Figure~\ref{fig:weights}
+and Figure~\ref{fig:weights_cut} are compared,
+it is clear that
 BMC algorithm now uses more information from
 morphological separation and \texttt{HB},
-while it uses considerably less information from TPC.
-%Another interesting pattern is that
-%\texttt{SOMc} and morphological separation seem complementary;
-%\texttt{SOMc} is weighted most strongly at fainter magnitudes,
-%while the morphological separation method
-%is weighted most strongly at brighter magnitudes.
-It is not surprising that the morphological separation method
+while it uses considerably less information from the 
+data-driven methods, \texttt{TPC} and \texttt{SOMc}.
+Not surprisingly, the morphological separation method
 performs best at bright magnitudes,
@@
 
+\textcolor{blue}{We present the star and galaxy purity values as a function of
+$i$-band magnitude for the differential counts
+in Figure~\ref{fig:purity_mag_cut}.
+Compred to Figure~\ref{fig:purity_mag},
+the purity of data-driven technique, TPC, suffers a significant
+decrease in purity.
+However, the purity of BMC does not show such a significant drop
+and decreases by only 2-5\%, and BMC still outperforms
+the morphological separation method at fainter magnitudes $i > 21$.
+while they are similar at bright magnitudes.
+As suggested by the weight maps in Figure~\ref{fig:weights_cut},
+BMC can accomplish this by shifting the weights assigned to
+each base classifier in different SOM cells,
+\ie assigning less weight to \texttt{TPC}
+and more weight to \texttt{HB} and morphological separation
+when the quality of training data is relatively poor.
+We can also see in the bottom two panels that
+the performance of morphological separation suffers
+at fainter magnitudes $i > 21$,
+while BMC shows consistently better performance over all magnitude ranges.}
+
 \begin{figure}
   \centering
-  \includegraphics[width=\linewidth]{figures/purity_mag_cut.eps}
+  \includegraphics[width=\linewidth]{figures/purity_mag_cut.pdf}
   \caption{Purity as a function of the $i$-band magnitude
@@
            in the VVDS W1 field with $i < 22$.
-           We repat in the top panel shows the histogram of objects
+           Top panel shows the histogram and KDE for number distribution
            in the test data set.
            The second panel shows the fraction of stars
-           in the VVDS W1 and W4 fields.
-           The error bars are Poisson.
+           in the VVDS W1 and W4 fields in green and blue, respectively.
            The bottom two panels compare
            the galaxy and star purity for morphological separation, TPC,
-           and BMC techniques as a function of magnitude.
-           The error bars for completeness and purity are not shown for
-           clarity, but they are approximately 10 percent.}
+           and BMC techniques as a function of magnitude.}
   \label{fig:purity_mag_cut}
@@
 
+\textcolor{blue}{In Figure~\ref{fig:purity_mag_cut_itnegrated}, we show
+the overall galaxy and star purity as a function of magnitude
+for the integrated counts.
+Even when \texttt{TPC} and \texttt{SOMc} are trained on heterogenous data,
+BMC is able to maintain a galaxy purity of 0.970
+up to $i \sim 24.5$, and it sill outperforms morphological separation
+at fainter magnitudes $i \gtrsim 21$.}
+
 \begin{figure}
   \centering
-  \includegraphics[width=\linewidth]{figures/purity_z_cut.eps}
+  \includegraphics[width=\linewidth]{figures/purity_mag_cut_integrated.pdf}
+  \caption{Purity as a function of the $i$-band magnitude
+           for the integrated counts.
+           The bottom two panels compare
+           the galaxy and star purity for BMC, TPC, and 
+           morphological separation as a function of magnitude.
+           Results for BMC, TPC, and morphological separation are in
+           red, green, and blue, respectively.
+           The error bars are Poisson.}
+  \label{fig:purity_mag_cut_integrated}
+\end{figure}
+
+We also show the star and galaxy purity values
+as a function of
+photometric redshift in Figure~\ref{fig:purity_z_cut},
+and as a function of $g-r$ in Figure~\ref{fig:purity_g_r_cut}.
+Compared to Figure~\ref{fig:purity_z} and \ref{fig:purity_g_r},
+where \texttt{TPC} and \texttt{SOMc} were trained
+on a high-quality training data,
+the performance of BMC is worse in some photo-$z$ and $g-r$ ranges.
+However, this drop in performance seems to be confined to
+only a small number of objects in particular parameter ranges,
+and BMC still outperforms both \texttt{TPC} and morphological separation
+for the majority of objects.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/purity_z_cut.pdf}
   \caption{The same figures as Figure~\ref{fig:purity_mag_cut},
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/purity_g-r_cut.eps}
+  \includegraphics[width=\linewidth]{figures/purity_g_r_cut.pdf}
   \caption{The same figure as Figure~\ref{purity_mag_cut},
@@
 
-We present the star and galaxy purity values
-of Table~\ref{table:metrics_vvds} as a function of
-$i$-band magnitude in Figure~\ref{fig:purity_mag_cut}.
-It is clear that BMC outperforms both
-morphological separation and TPC classification techniques
-over all magnitudes.
-The galaxy purity of BMC is significantly better
-than that of TPC at bright magnitudes, $i\lesssim21.5$,
-while they are similar at faint magnitudes.
-The star purity of BMC shows improvement over that of TPC
-at faint magnitudes,
-while they are similar at bright magnitudes.
-As suggested by the weight maps in Figure~\ref{fig:weights_cut},
-BMC can accomplish this by combining information from
-all base classifiers,
-\eg giving more weight to the morphological separation method
-at bright magnitudes.
-We can also see in the bottom two panels that
-the performance of morphological separation suffers
-at intermediate magnitudes $21\lesssim i \lesssim23$,
-while BMC shows consistently better performance
-over all magnitude ranges.
-
-We also show the star and galaxy purity values
-of Table~\ref{table:metrics_vvds} as a function of
-photometric redshift in Figure~\ref{fig:purity_z}.
-Photo-$z$ is estimated with
-the \texttt{BPZ} algorithm~\citep{benitez2000bayesian}
-and provided with
-the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
-It is again clear that BMC outperforms
-both individual classification techniques over all redshift range.
-
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/p_dist_all.eps}
-    \caption{Histogram of the posterior probability that
-             a source is a star for morphological separation (top),
-             TPC (middle), and BMC (bottom) techniques
-             for the entire training data
-             (the CFHTLenS data in the DEEP2, SDSS, VIPERS, and VVDS
-             fields as discussed in Section~\ref{section:rich_training}).}
-    \label{fig:p_dist_all}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/p_dist_cut.eps}
+The difference between the posterior probability distribution of
+TPC and that of BMC is now more pronounced in Figure~\ref{fig:p_dist_cut},
+where the $P\left(S\right)$ distribution of BMC for true galaxies
+falls off sharply at $P\left(S\right)\approx0.95$,
+while morphological separation and TPC techniques
+classify some true galaxies as stars with absolute certainty.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/p_dist_cut.pdf}
   \caption{The same figures as Figure~\ref{fig:p_dist_all},
@@
   \label{fig:p_dist_cut}
-  \end{minipage}
-\end{figure*}
-
-We show the distribution of posterior star probabilities
-in Figures~\ref{fig:p_dist_all} and \ref{fig:p_dist_cut}.
-The histogram of spectroscopic galaxies is in blue,
-and the histogram of true stars is in red.
-It is interesting that the BMC technique assigns
-a probability $P\left(S\right)$ greater than 0.6
-to almost no true galaxies in Figure~\ref{fig:p_dist_all},
-presumably because BMC utilizes information from
-different types of classification techniques.
-This pattern is more pronounced in Figure~\ref{fig:p_dist_cut},
-where the $P\left(S\right)$ distribution of BMC for true galaxies
-falls off sharply at $P\left(S\right)\approx0.95$,
-while morphological separation and TPC techniques
-classify some true galaxies as stars with absolute certainty.
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\linewidth]{figures/color_color.eps}
-  \caption{Color-color diagrams of sources with
-           $0.45<P\left(S\right)<0.55$ when the BMC technique
-           is applied to the poor training data.}
-  \label{fig:color_color}
-\end{figure*}
-
-To identify the regions in color-color space
-where our BMC technique struggles to classify stars and galaxies,
-we present the distribution over the color-color space for sources with
-$0.45 < P(S) < 0.55$ in Figure~\ref{fig:color_color}.
-Not surprisingly, these regions correspond to
-where there is a significant overlap between stars and galaxies
-in the color space.
-It may be possible to improve the performance of
-star-galaxy classification by obtaining more information
-through spectroscopic follow-up, manual study, or citizen science projects,
-or by developing a classification technique specifically designed
-for these regions of the color space and
-combining it with other techniques in a Bayesian framework.
+\end{figure}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=\linewidth]{figures/color_color.eps}
+%  \caption{Color-color diagrams of sources with
+%           $0.45<P\left(S\right)<0.55$ when the BMC technique
+%           is applied to the poor training data.}
+%  \label{fig:color_color}
+%\end{figure*}
+%
+%To identify the regions in color-color space
+%where our BMC technique struggles to classify stars and galaxies,
+%we present the distribution over the color-color space for sources with
+%$0.45 < P(S) < 0.55$ in Figure~\ref{fig:color_color}.
+%Not surprisingly, these regions correspond to
+%where there is a significant overlap between stars and galaxies
+%in the color space.
+%It may be possible to improve the performance of
+%star-galaxy classification by obtaining more information
+%through spectroscopic follow-up, manual study, or citizen science projects,
+%or by developing a classification technique specifically designed
+%for these regions of the color space and
+%combining it with other techniques in a Bayesian framework.
 
@@
 
-It is very costly in terms of telescope time to
-obtain a large sample spectroscopic observations
-down to the limiting magnitude of a photometric sample.
-Thus, we investigate the impact of training set quality
-by imposing various magnitude and $z$ cuts
-to decrease the size of the training set.
-
-In Figure.~\ref{fig:perform_mag_cut}, we present
+\textcolor{blue}{The BMC technique we have demonstrated in this work uses
+two training based algorithms as base classifiers.
+Ideally, the training data should mirror the entire parameter space
+occupied the data to be classified.
+Yet we have seen that the BMC technique do reliably
+extraplotate past the limits of the training data,
+even when some base classifiers are trained on a hetorogenous data
+with a magnitude cut imposed.
+In this section, we further investigate
+if and where BMC begins to break down
+by imposing various magnitude, photo-$z$, and color cuts
+to decrease the size of the training set.}
+
+\textcolor{blue}{In Figure.~\ref{fig:perform_mag_cut}, we present
 a visual comparison between different classification techniques
-when various magnitude cuts are applied.
-
-In Figure.~\ref{fig:perform_z_cut}, we also present
-the result of imposing various $z$ cuts.
-
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
+when various magnitude cuts are applied in the training data
+and the performance is measured on the same test set.
+The top panel shows that the size of training set
+decreases as we impose more restrictive magnitude cut.
+It is remarkable that BMC is able to reliably extrapolate
+past the training data and outperform \texttt{HB}, \texttt{TPC},
+and morphological separation in all performance metrics,
+even the demographics of training set does not accurately sample 
+the data to be classified and the size of training set is decreased.
+The performance is consistent until we impose a magnitude cut
+of $i < 20.5$ on the training data,
+beyond which point BMC finally performs worse than morphological separation.}
+
+\textcolor{blue}{Similarly, we also present impose various
+spectroscopic redshfit and color cuts in the training data
+in Figure~\ref{fig:perform_z_cut} and Figure~\ref{fig:perform_g_r_cut},
+resepctively.
+The area under the ROC curve measures the global performance,
+and BMC still outperforms \texttt{HB} and \texttt{TPC} globally in both cases,
+even when a very restrictive cuts are applied on the training set.
+Locally, BMC begins to perform worse than morphological separation
+when a conservative cut of $z_{\mathrm{spec}} < 0.6$ is imposed.
+A color cut seems to have more pronounced effect in the performance of BMC
+as shown in Figure~\ref{fig:perform_g_r_cut}.
+When the training data set is restricted to $g - r > 0.4$,
+BMC performs worse than morphological separation.}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{figures/perform_mag_cut.eps}
     \caption{The classification performance metrics for TPC, HB, morphology,
@@
     \label{fig:perform_mag_cut}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{figures/perform_z_cut.eps}
     \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
@@
     \label{fig:perform_z_cut}
-  \end{minipage}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{figures/perform_g_r.eps}
+    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
+      $g - r$ color cuts.}
+    \label{fig:perform_g_r_cut}
+\end{figure}
+
 \section{Conclusions}

### 2015-05-07 paper/main.tex
update paper.

--- 2015-04-15 paper/main.tex
@@
 \usepackage{url}
-\usepackage{float}
-
-\def\aap{{ A\&A}}
-\def\AaA{{ A\&A}}
+\usepackage{placeins}
+\usepackage{times}
+
+% By default the URLs are put in typewriter type in the body and the
+% bibliography of the document when using the \url command.  If you are
+% using many long URLs you may want to uncommennt the next line so they
+% are typeset a little smaller.
+\renewcommand{\UrlFont}{\small\tt}
+
+\def\aap{{A\&A}}
+\def\AaA{{A\&A}}
 \def\AJ{{AJ}}
@@
 
-\title{A Hybrid Ensemble Learning Approach to Star-galaxy Classification}
-
-\author[E. J. Kim et al.]{
-  Edward J. Kim\thanks{jkim575@illinois.edu}$^1$, Robert J. Brunner$^2$,
-  and Matias Carrasco Kind$^3$\\
+\pdfminorversion=5
+
+\title{A Hybrid Ensemble Learning Approach to Star-Galaxy Classification}
+
+\author[E. J. Kim, R. J. Brunner \& M. Carrasco Kind]{
+  Edward J.~Kim$^1$\thanks{jkim575@illinois.edu}, Robert J.~Brunner$^2$,
+  and Matias Carrasco Kind$^{2,3}$\\
 $^1$Department of Physics, University of Illinois, Urbana, IL 61801 USA\\
@@
 \begin{abstract}
-\textcolor{blue}{
-Star-galaxy separation in currently ongoing and future
-large photometric surveys, such as DES and LSST,
-will detect a vast number of unresolved galaxies at faint magnitudes.
+%Star-galaxy separation in currently ongoing and future
+%large photometric surveys, such as DES and LSST,
+%will detect a vast number of unresolved galaxies at faint magnitudes.
 There exist a variety of star-galaxy classification techniques,
-each with their own strenths and weaknesses.
-In this paper, we present a novel hybrid, ensemble classification
+each with their own strengths and weaknesses.
+In this paper, we present a novel meta-classification
 framework that combines and fully exploits different techniques
-to produce a more robust classification.
-To demonstrate this hybrid technique,
+to produce a more robust star-galaxy classification.
+To demonstrate this hybrid, ensemble approach,
 we combine a purely morphological classifier,
-a supervised machine learning technique based on random forest,
-an unsupervised method based on self-organizing maps,
-and a heirarchical Bayeisan template fitting algorithm.
-Any of these four methods could
-easily be replaced by a similar method
-or be extended to include other existing techniques.
-Testing with CFHTLenS data,
-we demonstrate that our combination technique improves
-the overall performance over any individual classification method.
-}
+a supervised machine learning method based on random forest,
+an unsupervised machine learning method based on self-organizing maps,
+and a Hierarchical Bayesian template fitting method.
+%Any of these four methods could
+%easily be replaced by a similar method
+%or be extended to include other existing techniques.
+Using data from the CFHTLenS survey,
+we consider different scenarios:
+when a high-quality training set is available with spectroscopic labels from
+DEEP2, SDSS, VIPERS, and VVDS, and
+when the demographics of sources in a low-quality training set
+do not match the demographics of objects in the test data set.
+We demonstrate that our Bayesian combination technique improves
+the overall performance over any individual classification method
+in these scenarios.
+Thus, strategies that combine the predictions of different classifiers
+may prove to be optimal in currently ongoing and forthcoming
+photometric surveys,
+such as the Dark Energy Survey and the Large Synoptic Survey Telescope.
 
@@
 \begin{keywords}
-\textcolor{red}{keywords}
+methods: data analysis -- methods: statistical -- surveys -- stars: statistics
+-- galaxies:statistics.
 \end{keywords}
@@
 
-The problem of star-galaxy classification is fundamental to astronomy
-and goes as far back as \cite{messier1781catalogue}.
+The problem of source classification is fundamental to astronomy
+and goes as far back as \citet{messier1781catalogue}.
 A variety of different strategies have been developed 
 to tackle this long-standing problem,
-and yet it remains an unsolved challenge.
+%and yet it remains an unsolved challenge.
+and yet there is no consensus on
+the optimal star-galaxy classification strategy.
 The most commonly used method to classify stars and galaxies
@@
 are provided along with the truth labels (\eg star or galaxy).
-\cite{odewahn1992automated} pioneered
+\citet{odewahn1992automated} pioneered
 the application of neural networks
@@
 Other successfully implemented examples include
-decision trees~\citep{weir1995automated, suchkov2005census, ball2006robust}
-and Support Vector Machines~\citep{Fadely2012}.
+decision trees~\citep{weir1995automated, suchkov2005census, ball2006robust,
+sevilla2015effect}
+and Support Vector Machines~\citep*{Fadely2012}.
 Unsupervised machine learning techniques
-are less common
-as they do not utilize the truth labels during the training process
+are less common,
+as they do not utilize the truth labels during the training process,
 and only the input attributes are used.
@@
 to the performance of the individual classification techniques.
+Finally, we outline our conclusions in Section~\ref{section:conclusions}.
 
@@
 four distinct star-galaxy classification techniques.
-The first method is
-a supervised machine learning technique named \texttt{TPC}
+The first method is a morphological separation method,
+which uses a hard cut in the half-light radius vs.\ magnitude plane.
+The second method is
+a supervised machine learning technique named TPC
 (Trees for Probabilistic Classification),
 which uses prediction trees and a random forest~\citep{carrascokind2013tpz}.
-The second method is
-an unsupervised machine learning technique named \texttt{SOMc},
-which uses self-organizing maps (SOM) and a random atlas
-to provide a classification~citep{carrascokind2014somz}.
 The third method is
+an unsupervised machine learning technique named SOMc,
+which uses self-organizing maps (SOMs) and a random atlas
+to provide a classification~\citep{carrascokind2014somz}.
+The fourth method is
 a Hierarchical Bayesian template fitting technique
-based on the work by \cite{Fadely2012},
+based on the work by \citet{Fadely2012},
 which fits SED templates from star and galaxy libraries
 to an observed set of measured flux values.
-The fourth method is a morphological separation method,
-which uses a hard cut in the half-light radius vs.\ magnitude plane.
 
@@
 we adopt a technique that is popular among the weak lensing 
-community~\citep{Kaiser1995}.
+community~\citep*{Kaiser1995}.
 As Figure~\ref{fig:morph} shows, there is a distinct locus
 produced by point sources in the half-light radius
-($r_h$; estimated by \textsc{SExtractor}'s \texttt{FLUX\_RADIUS} parameter)
+(estimated by \textsc{SExtractor}'s FLUX\_RADIUS parameter)
 vs.\ the $i$-band magnitude plane.
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{figures/morph.eps}
-  \caption{Half-light radius ($r_h$) vs.\ magnitude.}
+  \includegraphics[width=\columnwidth]{figures/morph.eps}
+  \caption{Half-light radius vs.\ magnitude.}
   \label{fig:morph}
@@
 A more informative approach is to provide probabilistic classifications.
-Although a recent work by \cite{henrion2011bayesian}
+Although a recent work by \citet{henrion2011bayesian}
 implemented a probabilistic classification using a Bayesian approach
@@
 In particular, we later show that
-the binary output (\ie 0 or 1) of cut-based methods
-can be transformed into a
-probability estimate by combining them
+the binary outputs (\ie 0 or 1) of cut-based methods
+can be transformed into
+probability estimates by combining them
 with the probability outputs from other
 probabilistic classification techniques,
-such as \texttt{TPC}, \texttt{SOMc}, and \texttt{HB}.
+such as TPC, SOMc, and HB.
 
@@
 
-\texttt{TPC} is a parallel, supervised machine learning algorithm
+TPC is a parallel, supervised machine learning algorithm
 that uses prediction trees and random forest 
 techniques~\citep{breiman1984classification, breiman2001random}
-to produce star-galaxy classification.
-\texttt{TPC} is a part of a publicly available software package called
-\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}
+to produce a star-galaxy classification.
+TPC is a part of a publicly available software package called
+\textsc{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}
 (Machine Learning for Photo-$z$).
-The full software package includes, among other capabilities,
-\texttt{TPZ}, a supervised photo-$z$ technique
-\citep[regression mode;][]{carrascokind2013tpz},
-\texttt{TPC}, a supervised star-galaxy classification technique
-(classification mode),
-\texttt{SOM}$z$, an unsupervised photo-$z$ technique
-\citep[regression mode;][]{carrascokind2014somz},
+The full software package includes:
+\textsc{TPZ}, a supervised photometric redshift (photo-$z$)
+estimation technique
+\citep[regression mode;][]{carrascokind2013tpz};
+TPC, a supervised star-galaxy classification technique
+(classification mode);
+\textsc{SOM}$z$, an unsupervised photo-$z$ technique
+\citep[regression mode;][]{carrascokind2014somz};
 and
-\texttt{SOMc}, an unsupervised star-galaxy classification technique
+SOMc, an unsupervised star-galaxy classification technique
 (classification mode).
 
-\texttt{TPC} uses classification trees,
+TPC uses classification trees,
 a type of prediction trees that are designed to
@@
 
-\begin{equation}
-  I_G \left(D_{\mathrm{node}}, X\right)
-  = I_d \left( D_{\mathrm{node}} \right)
-  - \sum_{x \in \mathrm{values}(X)}
-  \frac{|D_{\mathrm{node}, x}|}{|D_{\mathrm{node}}|}
-  I_d \left( D_{\mathrm{node}, x} \right),
-  \label{eq:information_gain}
+\begin{equation} \label{eq:information_gain}
+  I_G \left(D_{\rmn{node}}, X\right)
+  = I_d \left( D_{\rmn{node}} \right)
+  - \sum_{x \in \rmn{values}(X)}
+  \frac{|D_{\rmn{node}, x}|}{|D_{\rmn{node}}|}
+  I_d \left( D_{\rmn{node}, x} \right),
 \end{equation}
@@
 \noindent
-where $D_{\mathrm{node}}$ is the training data in a given node,
+where $D_{\rmn{node}}$ is the training data in a given node,
 $X$ is one of the possible dimensions (\eg magnitudes or colors)
@@
 $x$ are the possible values of a specific dimension $X$.
-$|D_{\mathrm{node}}|$ and $|D_{\mathrm{node}, x}|$ are the size of the total training data
+$|D_{\rmn{node}}|$ and $|D_{\rmn{node}, x}|$ are the size of the total training data
 and the number of objects in a given subset $x$ within the current node,
@@
 $I_d$ is the impurity degree index,
-and \texttt{TPC} can calculate $I_d$
+and TPC can calculate $I_d$
 from any of the three standard different impurity indices:
@@
 and \textit{classification error}.
-Here, we use the information entropy,
+In this work, we use the information entropy,
 which is defined similarly to the thermodynamic entropy:
@@
   I_d \left( D \right)
-  = - f_g \log_{2} f_g - (1 - f_g) \log_{2} f_g,
+  = - f_g \log_{2} f_g - \left(1 - f_g\right) \log_{2} \left(1 - f_g\right),
 \end{equation}
@@
 from the input training data
-by sampling repeatedly from the magnitudes using the magnitude errors.
+by sampling repeatedly from the magnitudes and colors
+using their measurement errors.
 We use these bootstrap samples to construct
@@
 
-We also use a cross validation technique called
+We also use a cross-validation technique called
 Out-of-Bag~\citep[OOB;][]{breiman1984classification, carrascokind2013tpz}.
-When a tree (or a map) is built in \texttt{TPC} (or \texttt{SOMc}),
+When a tree (or a map) is built in TPC (or SOMc),
 a fraction of the training data, usually one-third,
-is left out and not used in training the trees or maps.
+is left out and not used in training the trees (or maps).
 After a tree is constructed using two-thirds of the training data,
@@
 
-A Self-Organizing Map
-\citep[SOM;][]{kohonen1990self, kohonen2001self}
+A self-organizing map~\citep{kohonen1990self, kohonen2001self}
 is an unsupervised, artificial neural network algorithm
@@
 The output map is usually chosen to be two-dimensional
-for the visualization.
+so that the resulting map can be used for visualizing
+various properties of the input data.
 The differences between a SOM and other neural network algorithms are
@@
 In fact, a SOM can be viewed as a non-linear generalization
-of a principal component analysis (PCA) algorithm.
+of a principal component analysis (PCA) algorithm~\citep{yin2008self}.
 
@@
 
-In a similar approach to random forest in \texttt{TPZ} and \texttt{TPC},
-\texttt{SOM}$z$ uses a technique called \textit{random atlas}
+In a similar approach to random forest in TPZ and TPC,
+SOM$z$ uses a technique called \textit{random atlas}
 to provide photo-$z$ estimation \citep{carrascokind2014somz}.
@@
 Furthermore, we create random realizations of the training data
-by perturbing the input attributes by their measurement error.
+by perturbing the magnitudes and colors by their measurement errors.
 For each map, we can either use all available attributes,
@@
 applied to the classification problem,
-and we refer to it as \texttt{SOMc}
+and we refer to it as SOMc
 in order to differentiate it from 
@@
 
-One of the most important parameter in \texttt{SOMc}
+One of the most important parameter in SOMc
 is the topology of the two-dimensional SOM,
@@
 which is constructed by using
-\texttt{HEALPIX}~\citep{gorski2005healpix}.
-Furthermore, similar to \texttt{TPC},
+\textsc{HEALPIX}~\citep{gorski2005healpix}.
+Furthermore, similar to TPC,
 we use the OOB technique to make an unbiased estimation of errors.
@@
 and its application to the estimation of
-photo-$z$ probability distribution functions,
+photo-$z$ probability density functions (photo-$z$ PDFs),
 we refer the reader to \cite{carrascokind2014somz}.
@@
 
-\begin{equation}
-P \left( S | \mathbf{x}, \mathbf{\theta} \right)
-= P \left( \mathbf{x} | S, \mathbf{\theta} \right)
-P \left( S | \mathbf{\theta} \right),
+\begin{equation} \label{eq:overall_posterior}
+P \left( S | \bmath{x}, \bmath{\theta} \right)
+= P \left( \bmath{x} | S, \bmath{\theta} \right)
+P \left( S | \bmath{\theta} \right),
 \end{equation}
@@
 \noindent
-where $\mathbf{x}$ represents a given set of observed magnitudes,.
-We have also introduced the \textit{hyperparameter} $\mathbf{\theta}$,
+where $\bmath{x}$ represents a given set of observed magnitudes,.
+We have also introduced the \textit{hyperparameter} $\bmath{\theta}$,
 a nuisance parameter that characterizes our uncertainty
@@
 To compute the likelihood that a source is a star,
-we marginalize over all star and galaxy templates $\mathbf{T}$.
+we marginalize over all star and galaxy templates $\bmath{T}$.
 In a template-fitting approach,
 we marginalize by summing up
-the likelihood that a source has the set of magnitudes $\mathbf{x}$
+the likelihood that a source has the set of magnitudes $\bmath{x}$
 for a given star template
@@
 
-\begin{equation}
-  P \left(\mathbf{x} | S, \mathbf{\theta} \right)
-  = \sum_{t \in \mathbf{T}}
-  P \left(\mathbf{x} | S, t, \mathbf{\theta} \right)
-  P \left(t | S, \mathbf{\theta} \right).
-  \label{eq:marginalize_template}
+\begin{equation} \label{eq:marginalize_template}
+  P \left(\bmath{x} | S, \bmath{\theta} \right)
+  = \sum_{t \in \bmath{T}}
+  P \left(\bmath{x} | S, t, \bmath{\theta} \right)
+  P \left(t | S, \bmath{\theta} \right).
 \end{equation}
@@
 The likelihood of each template
-$P \left( \mathbf{x} | S, \mathbf{\theta} \right)$
+$P \left( \bmath{x} | S, \bmath{\theta} \right)$
 is itself marginalized over the uncertainty
@@
 requires that we specify the prior probability
-$P \left(t | S, \mathbf{\theta} \right)$
-that a source has spectral template $t$ (at a given redshift).
+$P \left(t | S, \bmath{\theta} \right)$
+that a source has a spectral template $t$ (at a given redshift).
 Thus, the probability that a source is a star (or a galaxy)
@@
 if a prior is used,
-or the likelihood itself if an uniformative prior is used.
+or the likelihood itself if an uninformative prior is used.
 In a Bayesian analysis, it is preferable to use a prior,
 which can be directly computed either from physical assumptions,
-from an empirical function calibrated by
-using a spectroscopic training sample,
 or from an empirical function calibrated by
-using machine learning techniques.
+using a spectroscopic training sample.
 In an HB approach, the entire sample of sources is used to
@@
 \begin{equation}
-\sum_{t \in \mathbf{T}}
-P \left(t | S, \mathbf{\theta} \right) = 1.
+\sum_{t \in \bmath{T}}
+P \left(t | S, \bmath{\theta} \right) = 1.
 \end{equation}
@@
 \noindent
+Similarly, we also parametrize the overall prior probability,
+$\left(S | \bmath{\theta}\right)$ in Equation~\ref{eq:overall_posterior},
+as a weight.
 These weights correspond to the hyperparameters,
@@
 the posterior probability distribution in the hyperparameter space.
-For the sampling, we use \texttt{emcee}, a Python implementation of the
+For the sampling, we use \textsc{emcee}, a Python implementation of the
 affine-invariant Markov Chain Monte Carlo (MCMC) ensemble sampler~\citep{Foreman-Mackey2013}.
@@
 For our stellar templates,
-we use the empirical SED library from \cite{pickles1998stellar}.
+we use the empirical SED library from \citet{pickles1998stellar}.
 The Pickles library consists of 131 stellar templates,
@@
 and luminosity classes at solar abundance,
-as well as metal-poor and metal-rich F-K dwarf 
-and G-K giant and supergiant stars.
-\textcolor{blue}{
+as well as metal-poor and metal-rich F--K dwarf 
+and G--K giant and supergiant stars.
 We supplement the stellar library with
-100 SEDs from \cite{chabrier2000evolutionary},
+100 SEDs from \citet{chabrier2000evolutionary},
 which include low mass stars and brown dwarfs
-with different $T_{eff}$ and surface gravities.
+with different $T_{\rmn{eff}}$ and surface gravities.
 We also include four white dwarf templates of
-\cite{bohlin1995white}.
-}
+\citet*{bohlin1995white}, for a total of 235 templates
+in our final stellar library.  
 For our galaxy templates,
-we use the four CWW spectra from \cite{coleman1980colors},
+we use four CWW spectra from \cite*{coleman1980colors},
 which include an Elliptical, an Sba, an Sbb,
 and an Irregular galaxy template.
-When extending an analysis to higher redshift,
+When extending an analysis to higher redshifts,
 the CWW library is often augmented with
 two star bursting galaxy templates from \cite{kinney1996template}.
-\textcolor{blue}{
 From the six original CWW and Kinney spectra,
-intermediate templates are created by interpolation
-for a total of 51 SEDs.
-}
+intermediate templates are created by interpolation,
+for a total of 51 SEDs in our final galaxy library.
 
@@
 individual star-galaxy classification techniques
-using ensemble learning techniques known as
-Bayesian Model Combination (BMC)
+using four combination techniques,
+including Bayesian Model Combination (BMC)
 and a variant of an ensemble learning technique known as stacking.
-The main idea behind ensemble learning is to weigh
+The main idea behind ensemble learning is to weight
 the predictions from individual models
@@
 Similarly, we can expect unsupervised approaches
-such as \texttt{SOM} or template fitting approaches 
+such as SOM or template fitting approaches 
 to generally perform better when training data
@@
 
-We therefore adopt the binning strategy similar to 
+We therefore adopt a binning strategy similar to 
 \cite{carrascokind2014exhausting}.
@@
 here we choose a 10$\times$10 rectangular topology to facilitate 
-visualization as shown in Figure.~\ref{fig:som_colors}.
+visualization as shown in Figure~\ref{fig:som_colors}.
 For all combination methods,
-we use only the OOB (cross validation) data contained in each cell
+we use only the OOB (cross-validation) data contained in each cell
 to compute the relative weights for the base classifiers.
@@
 To produce a single probability estimate for each source,
-one can take the mean, the median, or some other simple statistic.
+we could take the mean, the median, or some other simple statistic.
 With a sufficient number of maps,
 we find that there is usually negligible difference
-between taking the mean and taking the median,
-and so we use the median in the following sections.
+between taking the mean and taking the median, and
+use the median in the following sections.
 We note that it is also possible to establish confidence intervals
@@
 \begin{equation}
-P\left(S | \mathbf{x}, \mathbf{M} \right)
-= \sum_{i} P \left( S | \mathbf{x}, M_{i} \right),
+P\left(S | \bmath{x}, \bmath{M} \right)
+= \sum_{i} P \left( S | \bmath{x}, M_{i} \right),
 \end{equation}
@@
 \noindent
-where $\mathbf{M}$ is the set of models
-(\texttt{TPC}, \texttt{SOMc}, \texttt{HB}, and morphological separation
+where $\bmath{M}$ is the set of models
+(TPC, SOMc, HB, and morphological separation
 in our work).
@@
 by using the binning strategy to
-calculate the weighted average of each bin separately,
-rather than globally.
+calculate the weighted average of objects in each SOM cell
+separately for each map,
+and then combine the predictions from each map into a final prediction.
 
@@
 to choose the best model within each bin,
-and we use only that model within that specific bin
+and use only that model within that specific bin
 to make predictions for the test data.
@@
 
-\begin{equation}
-  \mathrm{MSE} = \frac{1}{N} \sum^{N - 1}_{i = 0}
+\begin{equation} \label{eq:mse}
+  \rmn{MSE} = \frac{1}{N} \sum^{N - 1}_{i = 0}
   \left( y_i - \hat{y}_i \right)^2,
-  \label{eq:mse}
 \end{equation}
@@
 Although any arbitrary algorithm
-can theoreticaly be used as a meta-classifier,
+can theoretically be used as a meta-classifier,
 a logistic regression or a linear regression is often used
@@
 we use a single-layer multi-response linear regression algorithm,
-which often shows the best performance~\citep{breiman1996stacked, ting1999issues}.
+which often shows the best performance~\citep{breiman1996stacked,
+ting1999issues}. 
+This algorithm
+is a variant of the least-square regression algorithm,
+where a linear regression model is constructed for each class.
 
@@
 
-We also use a model combination technique \citep{Monteith2011} known as
-Bayesian Model Combination (BMC).
-BMC uses Bayesian principles to generate an ensemble combination of
-different classifiers~\citep{Monteith2011}.
+We also use a model combination technique known as
+Bayesian Model Combination~\citep[BMC;][]{Monteith2011}, which
+uses Bayesian principles to generate an ensemble combination of
+different classifiers.
 The posterior probability that a source is a star is given by
 
-\begin{equation}
-  P \left(S | \mathbf{x}, \mathbf{D}, \mathbf{M}, \mathbf{E} \right)
-  = \sum_{e \in \mathbf{E}} P \left(S | \mathbf{x}, \mathbf{M}, e \right)
-  P \left(e | \mathbf{D} \right),
-  \label{eq:p_star_bmc}
+\begin{equation} \label{eq:p_star_bmc}
+  P \left(S | \bmath{x}, \bmath{D}, \bmath{M}, \bmath{E} \right)
+  = \sum_{e \in \bmath{E}} P \left(S | \bmath{x}, \bmath{M}, e \right)
+  P \left(e | \bmath{D} \right),
 \end{equation}
@@
 \noindent
-where $\mathbf{D}$ is the data set,
-and $e$ is an element in the ensemble space $\mathbf{E}$ of possible model combinations.
-By Bayes' Theorem, the posterior probability of $e$ given $\mathbf{D}$ is given by
-
-\begin{equation}
-  P \left(e | \mathbf{D} \right)
-  = \frac{P \left(e \right)}{P \left(\mathbf{D} \right)}
-  \prod_{d \in \mathbf{D}} P \left( d | e \right)
-  \propto P \left(e\right) \prod_{d \in \mathbf{D}} P \left(d | e \right).
-  \label{eq:p_ensemble}
+where $\bmath{D}$ is the data set,
+and $e$ is an element in the ensemble space $\bmath{E}$ of possible model combinations.
+By Bayes' Theorem, the posterior probability of $e$ given $\bmath{D}$ is given by
+
+\begin{equation} \label{eq:p_ensemble}
+  P \left(e | \bmath{D} \right)
+  = \frac{P \left(e \right)}{P \left(\bmath{D} \right)}
+  \prod_{d \in \bmath{D}} P \left( d | e \right)
+  \propto P \left(e\right) \prod_{d \in \bmath{D}} P \left(d | e \right).
 \end{equation}
@@
 \noindent
-Here, the product of $P\left(d \mid e\right)$ is
-over all individual data $d$ in the training data $\mathbf{D}$,
-and $P\left(\mathbf{D}\right)$ is merely a normalization factor
+Here, $P\left(e \right)$ is the prior probability of $e$,
+which we assume to be uniform.
+The product of $P\left(d | e\right)$ is
+over all individual data $d$ in the training data $\bmath{D}$,
+and $P\left(\bmath{D}\right)$ is merely a normalization factor
 and not important.
@@
 
-\begin{equation}
-  P \left( e | \mathbf{D} \right) \propto 
-  P \left( e \right) \left(1 - \epsilon \right)^{N_s + N_g}
+\begin{equation} \label{eq:p_ensemble_discrete}
+  P \left( e | \bmath{D} \right) \propto 
+  P \left( e \right) \left(1 - \epsilon \right)^{N - M_s + M_g}
   \left( \epsilon \right)^{M_s + M_g}.
@@
 
-\begin{equation}
-  P \left( e | \mathbf{D} \right) \propto 
+\begin{equation} \label{eq:p_ensemble_prob}
+  P \left( e | \bmath{D} \right) \propto 
   P \left( e \right) \prod_{i=0}^{N-1}
@@
 
-Although the space $\mathbf{E}$ of potential model combinations
+Although the space $\bmath{E}$ of potential model combinations
 is in principle infinite,
@@
 is obtained by sampling from a Dirichlet distribution.
-We first set all alpha values of a Dirichlet distribution.
+We first set all alpha values of a Dirichlet distribution to unity.
 We then sample this distribution $q$ times
@@
 we assume a uniform prior and
-calculate $P\left(e|\mathbf{D}\right)$ using Equation~\ref{eq:p_ensemble}.
-We select the combination with the highest $P\left(e|\mathbf{D}\right)$,
+calculate $P\left(e|\bmath{D}\right)$ using
+Equation~\ref{eq:p_ensemble_discrete} or \ref{eq:p_ensemble_prob}.
+We select the combination with the highest $P\left(e|\bmath{D}\right)$,
 and update the alpha values by
@@
 to the current alpha values.
-The next $q$ weight sets of weights are drawn
+The next $q$ sets of weights are drawn
 using the updated alpha values.
@@
 In this paper, we use a $q$ value of three,
-and 1,000 model combinations were considered.
+and 1,000 model combinations are considered.
 
@@
 We however note that introducing randomness
-though the construction of different SOM representations
+though the construction of $N_R \times N_M$ different SOM representations
 does not show significant
@@
 that has already been introduced by 
-the sampling from the Dirichlet distribution.
+sampling from the Dirichlet distribution.
 Thus, our BMC technique uses one SOM,
@@
 \cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
-{Heymans2012,Erben2013,Hildebrandt2012}.
-This catalog consits of more than twenty five million objects
+{heymans2012cfhtlens,erben2013cfhtlens,hildebrandt2012cfhtlens}.
+This catalog consists of more than twenty five million objects
 with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
@@
 the Deep Extragalactic Evolutionary Probe Phase 2~
-\citep[DEEP2;][]{Davis2003,Newman2013a},
+\citep[DEEP2;][]{davis2003science,newman2013deep2},
 the Sloan Digital Sky Survey Data Release 10~\citep[SDSS-DR10]{Ahn2014},
@@
 Very Large Telescope (VLT) Deep Survey~
-\citep[VVDS;][]{LeFevre2005,Garilli2008}, and
-the VIMOS Public Extragalactic Redshift Survey~\citep[VIPERS;][]{Garilli2014},
+\citep[VVDS;][]{le2005vimos,garilli2008vimos}, and
+the VIMOS Public Extragalactic Redshift
+Survey~\citep[VIPERS;][]{garilli2014vimos},
 In the end, we have 8,545 stars and 57,843 galaxies available
 for the training and testing processes.
+We randomly select 13,278 objects for the blind testing set,
+and use the remainder for training and cross-validation.
 
@@
 particular properties of the CFHTLenS data.
-For example, \cite{Hildebrandt2012} suggest that
+For example, \cite{hildebrandt2012cfhtlens} suggest that
 all objects with $i > 23$ in the CFHTLenS dataset
 may be classified as galaxies
-without significant incompleteness in the galaxy sample.
+without significant incompleteness and contamination in the galaxy sample.
 Although this approach works because
@@
 it can also be converted into a class label
-by using a threshold (probability cut).
+by using a threshold (a probability cut).
 The simplest way to choose the threshold is to set it to a fixed value,
-\eg $p_\mathrm{cut} = 0.5$.
+\eg $p_\rmn{cut} = 0.5$.
 This is, in fact, what is often done
@@
 
+\subsubsection{Receiver Operating Characteristic Curve}
+
 When we have no information about the operating condition
@@
 the Receiver Operating Characteristic (ROC) curve
-\citep{swets2000better}.
+\citep*{swets2000better}.
 An ROC curve is a graphical plot that illustrates the true positive rate
 versus the false positive rate of a binary classifier
-as its classification threshold (probability cut) is varied.
+as its classification threshold is varied.
 The Area Under the Curve (AUC) summarizes the curve information
@@
 
+\subsubsection{Completeness and Purity}
+
 In astronomical applications,
 the operating condition usually translates to
-the completeness and purity requirements of the star/galaxy sample,
+the completeness and purity requirements of the star or galaxy sample.
 We define the galaxy \textit{completeness}
@@
 To compare the performance of probabilistic classification techniques
-with that of the morphological separation,
+with that of morphological separation,
 which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
@@
 we adjust the threshold of probabilistic classifiers
-until the galaxy completeness matches that of the morphological classifier
-to compute $p_{g}$ at $c_g=0.9964$.
+until the galaxy completeness $c_g$ matches
+that of morphological separtion
+to compute the galaxy purity $p_{g}$ at $c_g=0.9964$.
 Similarly, the star purity $p_{s}$ at $c_{s}=0.7145$
-is computed by adjusting the threshold values
+is computed by adjusting the threshold
 until the star completeness of each classifier is equal to
-that of the morphological classifier.
+that of morphological separation.
 
@@
 Thus, we compute $p_{g}$ at $c_g=0.960$
-and $c_{s}$ at $p_s=0.250$.
+and $p_{s}$ at $c_s=0.250$.
 We also use the MSE defined in Equation~\ref{eq:mse}
@@
   $p_s$ & Star purity \\
-  $p_g(c_g=x)$ & Galaxy purity at $x$ galaxy purity \\
-  $p_s(c_s=x)$ & Star purity at $x$ star purity \\
+  $p_g(c_g=x)$ & Galaxy purity at $x$ galaxy completeness \\
+  $p_s(c_s=x)$ & Star purity at $x$ star completeness \\
   \end{tabular}
@@
 classification techniques.
-Given the wealth of training data, it is not surprising that 
-our supervised machine learning tehcnique \texttt{TPC}
-outperforms other unsupervised technique.
-\texttt{TPC} is thus shown in the first row as the benchmark.
+Given a high-quality training data, it is not surprising that 
+our supervised machine learning technique TPC
+outperforms other unsupervised techniques.
+TPC is thus shown in the first row as the benchmark.
 
 The simplest of the combination techniques, WA and BoM,
-generally do not perform better than \texttt{TPC}.
+generally do not perform better than TPC.
 It is also interesting that,
@@
 within each bin,
-BoM almost always chooses \texttt{TPC} as the best model,
-and therefore gives the same performance as \texttt{TPC} in the end.
+BoM almost always chooses TPC as the best model in all bins,
+and therefore gives the same performance as TPC in the end.
 However, our BMC and stacking techniques have a similar performance
-and often outperform \texttt{TPC}.
-\textcolor{blue}{
-Although \texttt{TPC} shows the best performance as measured by the AUC,
+and often outperform TPC.
+Although TPC shows the best performance as measured by the AUC,
 BMC shows the best performance in all other metrics.
-}
-
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/som_colors.eps}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/som_colors.pdf}
   \caption{A two-dimensional 10$\times$10 SOM representation
@@
            the fraction of true stars in each cell (top right),
-             and the mean values of $u-g$ and $g-r$
-             for the cross validation data.}
+           and the mean values of $u-g$ (bottom left) and $g-r$ (bottom right)
+           for the cross-validation data.}
   \label{fig:som_colors}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/weights.eps}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/weights.pdf}
   \caption{A two-dimensional 10$\times$10 SOM representation
@@
   \label{fig:weights}
-  \end{minipage}
-\end{figure*}
+\end{figure}
 
 In Figure~\ref{fig:som_colors}, we show
-the mean CFHTLenS $i$-band magnitude in the top left panel
-\textcolor{blue}{
+the mean CFHTLenS $i$-band magnitude in the top left panel,
 and the fraction of stars in each cell in the top right panel.
-}
-The bottom two panels show the mean $u-g$ and  $g-r$ colors.
-These two-dimensional maps clearly shows
+The bottom two panels show the mean $u-g$ and  $g-r$ colors in each cell.
+These two-dimensional maps clearly show
 the ability of the SOM to preserve relationships between sources
 when it projects the full nine-dimensional space to the two-dimensional map.
-The SOM mapping is a non-linear representation of all magnitudes and colors,
-and thus these SOM maps should only be used to provide guidance.
-
-We can also use the same SOM to determine the relative weights for 
-the four individual classification methods for each cell.
+We note that these SOM maps should only be used to provide guidance,
+as the SOM mapping is a non-linear representation of all magnitudes and colors.
+
+We can also use the same SOM from Figure~\ref{fig:som_colors}
+to determine the relative weights for 
+the four individual classification methods in each cell.
 We present the four weight maps for the BMC technique 
 in Figure~\ref{fig:weights}.
-In these maps, a redder color indicates a higher weight,
-or equivalently that the corresponding weight performs better in that region.
+In these maps, a darker color indicates a higher weight,
+or equivalently that the corresponding classifier
+performs better in that region.
 These weight maps demonstrate the variation in
@@
 the two-dimensional parameter space defined by the SOM.
+Furthermore, since the maps in Figure~\ref{fig:som_colors}
+and \ref{fig:weights} are constructed using the same SOM,
+we can determine the region in the parameter space
+where each individual technique performs better or worse.
 Not surprisingly, the morphological separation
-performs best in the top left corner of the $i$-band magnitude map,
-which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
-On the other hand, \texttt{TPC} seems to perform best
+performs best in the top left corner of the weight map
+in Figure~\ref{fig:weights},
+which corresponds to the brightest CFHTLenS magnitudes $i \la 20$
+in the $i$-band magnitude map of Figure~\ref{fig:som_colors}.
+It is also clear that the SOM cells where the morphological
+separation performs best have higher stellar fraction than
+the other cells. 
+On the other hand, TPC seems to perform best
 in the region that corresponds to intermediate magnitudes
-$20\gtrsim i \lesssim22.5$ and $1.5<u-g<3.0$.
-Our unsupervised learning method \texttt{SOMc}
-performs relatively better at fainter magnitudes $i>21.5$
-with $0<u-g<0.5$ and $0<g-r<0.5$.
-\textcolor{blue}{
-Although \texttt{HB} shows the worst performance
-when there exists ample training data,
-BMC still utilizes information from \texttt{HB}
-to produce an overall improvement in performance,
-especially at intermediate magnitudes $20\gtrsim i \lesssim22$.
-}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/purity_mag.eps}
-  \caption{Purity of morphological separation,
-           TPC, and BMC techniques
-           as a function of the $i$-band magnitude.
-           The top panel shows the histogram of objects in the test data set.
-           The second panel shows the fraction of stars
-           as a function of magnitude. The error bars are Poisson.
-           The bottom two panels compare
-           the galaxy and star purity for morphological separation, TPC,
-           and BMC techniques as a function of magnitude.
-           The error bars for completeness and purity are not shown for
-           clarity, but they are approximately 10 percent.}
-  \label{fig:purity_mag}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/purity_z.eps}
-  \caption{The same figures as Figure~\ref{fig:purity_mag},
-           but as a function of photometric redshift estimate
-           (photo-$z$).}
-  \label{fig:purity_z}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/purity_g-r.eps}
-  \caption{The same figure as Figure~\ref{purity_mag},
-           but as a function of $g-r$ color.}
-  \label{fig:purity_g-r}
-\end{figure}
-
-\textcolor{blue}{In Figure~\ref{fig:purity_mag}, 
+$20\la i \la22.5$ and $1.5 \la u-g \la 3.0$.
+Our unsupervised learning method SOMc
+performs relatively better at fainter magnitudes $i \ga 21.5$
+with $0 \la u-g \la 0.5$ and $0 \la g-r \la 0.5$.
+Although HB shows the worst performance
+when there exists a high-quality training data set,
+BMC still utilizes information from HB,
+especially at intermediate magnitudes $20\la i \la22$.
+Another interesting pattern is that
+the four techniques seem complementary,
+and they are weighted most strongly in different regions
+of the SOM representation.
+
+In Figure~\ref{fig:purity_mag}, 
 we compare the star and galaxy purity values
-for morphological separation, TPC, and BMC techniques
-as a function of $i$-band magnitude.
+for BMC, TPC, and morphological separation
+as functions of $i$-band magnitude for the differential counts.
+We use the kernel density estimation \cite[KDE;][]{silverman1986density}
+with the Gaussian kernel to smooth the fluctuations in the distribution.
 Although morphological separation shows a slightly better performance
-at bright magnitudes $i < 21$,
-BMC outperforms both morphological separation and TPC
-at faint magnitudes $i > 21$.
-As shown in Figure.~\ref{fig:purity_mag},
-the histogram of distribution of objects peaks at $i \sim 22$,
-and BMC therefore outperforms the morphological separation technique
+in galaxy purity at bright magnitudes $i \la 20$,
+BMC outperforms both TPC and morphological separation
+at faint magnitudes $i \ga 21$.
+As the top panel shows,
+the number count distribution peaks at $i \sim 22$,
+and BMC therefore outperforms both TPC and morphological separation
 for the majority of objects.
@@
 \eg giving more weight to the morphological separation method
-at bright magnitudes.}
-
-\textcolor{blue}{We also show the star and galaxy purity values
-of Table~\ref{table:metrics_all} as a function of
-photometric redshift in Figure~\ref{fig:purity_z}.
+at bright magnitudes.
+The bottom panels shows that
+the star purity of morphological separation drops to $p_s < 0.8$
+at fainter magnitudes $i > 21$.
+This is expected, as our crude morphological separation classifies
+every object as a galaxy beyond $i > 21$, and
+purity measures the number of true stars classified as stars.
+It is again clear that BMC outperforms both TPC and morphological separation
+in star purity values over all magnitudes.
+
+In Figure~\ref{fig:purity_mag_integrated}, we show
+the overall galaxy and star purity values as functions of magnitude
+for the integrated counts.
+Although morphological separation performs better than TPC
+at bright magnitudes, its purity values decrease as 
+the magnitudes become fainter, and
+TPC eventually outperforms morphological separation by 1--2\% at
+$i > 21$.
+BMC clearly outperforms TPC, the benchmark base classifier,
+and it maintains the overall galaxy purity of 0.980
+up to $i \sim 24.5$.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/purity_mag.pdf}
+  \caption{Purity as a function of the $i$-band magnitude
+           as estimated by the kernel density estimation (KDE) method
+           for the differential counts.
+           The top panel shows the histogram with a bin size of 0.1 mag
+           and the KDE for objects in the test set.
+           The second panel shows the fraction of stars estimated by KDE
+           as a function of magnitude.
+           The bottom two panels compare
+           the galaxy and star purity values for BMC, TPC, and 
+           morphological separation as functions of magnitude.
+           Results for BMC, TPC, and morphological separation are in
+           blue, green, and red, respectively.
+           The $1 \sigma$ confidence bands are estaimted by
+           bootstrap sampling.}
+  \label{fig:purity_mag}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/purity_mag_integrated.pdf}
+  \caption{Purity as a function of the $i$-band magnitude
+           for the integrated counts.
+           The upper panel compares
+           the galaxy purity values for BMC (blue solid line),
+           TPC (green dashed line), and 
+           morphological separation (red dashed line).
+           The lower panel compares the star purity.
+           The $1 \sigma$ error bars are computed following the method
+           of \citet{paterno2004calculating} to avoid the unphysical
+           errors of binomial or Poisson statistics.}
+  \label{fig:purity_mag_integrated}
+\end{figure}
+
+We also show the star and galaxy purity values as a function of
+photometric redshift estimate in Figure~\ref{fig:purity_z}.
 Photo-$z$ is estimated with
-the \texttt{BPZ} algorithm~\citep{benitez2000bayesian}
+the \textsc{BPZ} algorithm~\citep{benitez2000bayesian}
 and provided with
-the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
-Although the morphological separation method outperformed BMC
-at bright magnitudes in Figure.~\ref{fig:purity_mag},
+the CFHTLenS photometric redshift catalogue~\citep{hildebrandt2012cfhtlens}.
+The advantage of BMC over TPC and morphological separation is
+now more pronounced in Figure~\ref{fig:purity_z}.
+Although the morphological separation method outperforms BMC
+at bright magnitudes in Figure~\ref{fig:purity_mag},
 it is clear that BMC outperforms
-both morphological separation and TPC over all redshift range.}
-
-\textcolor{blue}{In Figure.~\ref{fig:purity_g-r},
-we present how the star and galaxy purity values vary
+both TPC and morphological separation over all redshift range.
+We also present in Figure~\ref{fig:purity_g_r}
+how the star and galaxy purity values vary
 as a function of $g-r$ color.
 It is again clear that BMC outperforms
-both morphological separation and TPC over all $g-r$ color range.  }
+both TPC and morphological separation over all $g-r$ color range.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/purity_z.pdf}
+  \caption{Similar to Figure~\ref{fig:purity_mag}
+           but as a function of photo-$z$.
+           The bin size of histogram in the top panel is 0.02.}
+  \label{fig:purity_z}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/purity_g_r.pdf}
+  \caption{Similar to Figure~\ref{fig:purity_mag}
+           but as a function of $g-r$ color.
+           The bin size of histogram in the top panel is 0.05.}
+  \label{fig:purity_g_r}
+\end{figure}
+
+In Figure~\ref{fig:p_dist_all},
+we show the distribution of $P(S)$,
+the posterior probability that an object is a star,
+for BMC, TPC, and morphological separation.
+It is interesting that the BMC technique assigns
+a posterior star probability $P(S) \la 0.3$
+to significantly more true galaxies than TPC,
+and a probability $P(S) \ga 0.8$ to significantly fewer true galaxies.
+By utilizing information from different types of classification techniques
+in different parts of the parameter space,
+BMC becomes more certain that an object is a star or a galaxy,
+resulting in improvement of overall performance.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/p_dist.pdf}
+  \caption{Histogram of the posterior probability that
+           a source is a star for morphological separation (top),
+           TPC (middle), and BMC (bottom)
+           for a high-quality training data set.
+           The true galaxies are in green, and true stars are in blue.
+           The bin size is 0.05.}
+  \label{fig:p_dist_all}
+\end{figure}
 
@@
 It is very costly in terms of telescope time to
-obtain a large sample spectroscopic observations
+obtain a large sample of spectroscopic observations
 down to the limiting magnitude of a photometric sample.
 Thus, we investigate the impact of training set quality
-by analyzing in detail a more realistic case
-where the training data is available
-only for a small number of objects at bright magnitudes.
+by considering a more realistic case
+where the training data set is available
+only for a small number of objects with bright magnitudes.
 To emulate this scenario,
@@
 and impose a magnitude cut of $i < 22.0$ in the training data,
-leaving us a training set with 1,365 objects.
+leaving us a training set with only 1,365 objects.
 We apply the same four star-galaxy classification techniques
-and four different combination methods,
-and test them on the same test data.
-Since the demographics of sources in the test data set
-does not show the same distribution as the training data set,
-this also serves as a test of the efficacy of heterogenous training data.
+and four combination methods,
+and measure the performance of each technique on the same test data set
+from Section~\ref{section:rich_training}.
+As the top two panels of Figures~\ref{fig:purity_mag_cut},
+\ref{fig:purity_z_cut}, and \ref{fig:purity_g_r_cut} show,
+the demographics of objects in the training set
+are different from the distribution of sources in the test set.
+Thus, this also serves as a test of the efficacy of heterogeneous training.
 
@@
            and the four different classification combination methods
-           as applied to the CFHTLenS data in the VVDS W1 field
-           with a magnitude cut of $i < 22.0$.
+           for when the training data set consists of
+           only the sources that are in CFHTLS W1 field,
+           have spectroscopic labels available from VVDS,
+           and have $i < 22$.
            The definition of the metrics is summarized in
@@
 Overall, the results obtained for the reduced data set are remarkable.
-With poor training data, our data-driven methods,
-\texttt{TPC} and \texttt{SOMc},
-suffer a significant decrease in performance,
-although they still in general perform better than 
-the template fitting algorithm
-\texttt{HB}.
-The performance of our morphological separation method
-does not depend on the training data,
-and shows the best result in the MSE metric.
-It also outperforms other base classifiers
-in galaxy purity at galaxy completeness of 0.9964
-and in star purity at star completeness of 0.7145.
-
+With a smaller training set, our training based methods, TPC and SOMc,
+suffer a significant decrease in performance.
+The performance of morphological separation and HB
+is essentially unchanged from Table~\ref{table:metrics_all}
+as they do not depend on the training data.
 Without sufficient training data,
-the advantage of classification combination techniques
+the advantage of combining the predictions of different classifiers
 is more obvious.
 Even WA, the simplest of combination techniques, outperforms
-all individual classification techniques.
-Interestingly, although TPC does not perform better
-than morphological separation,
-BoM still chooses TPC as the best model within each SOM cell
-and shows exactly the same performance metrics as TPC.
-Stacking and BMC techniques still outperform
-all individual classification techniques,
-and BMC's improvement in performance over TPC or mophological separation
-is impressive.
+all individual classification techniques in four metrics,
+AUC, $p_s$ at $c_s=0.7145$, $p_g$ at $c_g=0.9600$, and $p_s$ at $c_s=0.2500$.
+%Interestingly, although TPC does not perform better
+%than morphological separation,
+%BoM still chooses TPC as the best model within each SOM cell
+%and shows exactly the same performance metrics as TPC.
+%Stacking and BMC techniques still outperform
+%all individual classification techniques,
+Alhtough BoM always chooses TPC as the best model
+when there is a high-quality training set,
+it now chooses various methods and outperforms all base classifiers.
+While the performance of the stacking technique is comparable
+to that of BMC in Section~\ref{section:rich_training},
+stacking now fails to outperform morphological separation.
+BMC shows an impressive performance and 
+outperforms all other classification techniques in all six metrics.
 Overall, the improvements are small but still significant
@@
 %  \centering
-%  \includegraphics[width=\linewidth]{figures/roc_cut.eps}
+%  \includegraphics[width=\columnwidth]{figures/roc_cut.eps}
 %  \caption{A magnified view of the ROC curve of
@@
 
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/som_colors_cut.eps}
-    \caption{The same figure as Figure~\ref{fig:som_colors}
-        but for the BMC technique applied to the four individual methods
-        trained with the CFHTLenS data in the VVDS W1 field
-        with $i < 22$.}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/som_colors_cut.pdf}
+  \caption{Similar to Figure~\ref{fig:som_colors}
+           but for the reduced training data set. }
   \label{fig:som_colors_cut}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/weights_cut.eps}
-    \caption{The same figure as Figure~\ref{fig:weights}
-        but for the BMC technique applied to the four invidiual methods
-        trained with the CFHTLenS data in the VVDS W1 field
-        with $i < 22$.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/weights_cut.pdf}
+  \caption{Similar to Figure~\ref{fig:weights}
+           but for the reduced training data set. }
   \label{fig:weights_cut}
-  \end{minipage}
-\end{figure*}
+\end{figure}
 
@@
 two-dimensional weight map defined by the SOM.
-%It seems that \texttt{TPC} performs well
+%It seems that TPC performs well
 %in most of the parameter space.
-\textcolor{blue}{When the quality of training data is relatively poor,
-the performance of data driven algorithms will decrease,
-while the performance of template fitting algorithms such as \texttt{HB}
+When the quality of training data is relatively poor,
+the performance of training based algorithms will decrease,
+while the performance of template fitting algorithms
+or morphological separation methods
 is independent of training data.
-Thus, comparing the weight maps of Figure~\ref{fig:weights}
-and Figure~\ref{fig:weights_cut},
-it is perhaps not surprising that
-BMC algorithm now uses more information from
-morphological separation and \texttt{HB},
-while it uses considerably less information from TPC.
-%Another interesting pattern is that
-%\texttt{SOMc} and morphological separation seem complementary;
-%\texttt{SOMc} is weighted most strongly at fainter magnitudes,
-%while the morphological separation method
-%is weighted most strongly at brighter magnitudes.
-It is not surprising that the morphological separation method
+Thus, when the weight maps of Figure~\ref{fig:weights}
+and Figure~\ref{fig:weights_cut} are visually compared,
+it is clear that
+the BMC algorithm now uses more information from
+morphological separation and HB,
+while it uses considerably less information from 
+our training based algorithms, TPC and SOMc.
+Not surprisingly, the morphological separation method
 performs best at bright magnitudes,
-and BMC assigns more weight to \texttt{HB} at fainter magnitudes.}
+and BMC assigns more weight to HB at fainter magnitudes.
+
+We present the star and galaxy purity values as functions of
+$i$-band magnitude for the differential counts
+in Figure~\ref{fig:purity_mag_cut}.
+The normalized density distribution as a function of magnitude
+in the top panel
+and the stellar distribution in the second panel
+clearly show that the demographics of the training set and 
+that of the test set are different.
+Since the training set is cut at $i < 22$
+the density distribution falls off sharply around $i \sim 22$
+and has a higher fraction of stars than the test set.
+Compared to Figure~\ref{fig:purity_mag},
+the purity of TPC suffers a significant decrease in purity.
+However, the purity of BMC does not show such a significant drop
+and decreases by only 2--5\%.
+As suggested by the weight maps in Figure~\ref{fig:weights_cut},
+BMC can accomplish this by shifting the relative weights assigned to
+each base classifier in different SOM cells.
+As the quality of training set worsens,
+BMC assigns less weight to training based methods
+and more weight to HB and morphological separation.
+
+In Figure~\ref{fig:purity_mag_cut_integrated}, we show
+the overall galaxy and star purity values as functions of magnitude
+for the integrated counts.
+Compared to Figure~\ref{fig:purity_mag_integrated},
+the drop in the performance of TPC is clear.
+However, even when some classifers have been trained on
+a significantly reduced training set,
+BMC maintains a galaxy purity of 0.970 and a star purity of 1.0
+up to $i \sim 24.5$, and it sill outperforms morphological separation
+at fainter magnitudes $i \ga 21$.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/purity_mag_cut.eps}
+  \includegraphics[width=\columnwidth]{figures/purity_mag_cut.pdf}
   \caption{Purity as a function of the $i$-band magnitude
-           for morphological separation, TPC, and BMC techniques
-           in the VVDS W1 field with $i < 22$.
-           We repat in the top panel shows the histogram of objects
-           in the test data set.
+           for the reduced training data set.
+           Top panel shows the histograms and KDEs
+           for the number count distribution for
+           the training (blue) and test (green) data set.
            The second panel shows the fraction of stars
-           in the VVDS W1 and W4 fields.
-           The error bars are Poisson.
+           in the training and test data set in blue and green, respectively.
            The bottom two panels compare
-           the galaxy and star purity for morphological separation, TPC,
-           and BMC techniques as a function of magnitude.
-           The error bars for completeness and purity are not shown for
-           clarity, but they are approximately 10 percent.}
+           the galaxy and star purity values for BMC, TPC, 
+           and morphological separation as functions of $i$-band magnitude.}
   \label{fig:purity_mag_cut}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/purity_z_cut.eps}
-  \caption{The same figures as Figure~\ref{fig:purity_mag_cut},
+  \includegraphics[width=\columnwidth]{figures/purity_mag_cut_integrated.pdf}
+  \caption{Similar to Figure~\ref{fig:purity_mag_integrated}
+           but for the reduced training data set.}
+  \label{fig:purity_mag_cut_integrated}
+\end{figure}
+
+We also show the star and galaxy purity values as functions of
+photo-$z$ in Figure~\ref{fig:purity_z_cut}
+and as functions of $g-r$ in Figure~\ref{fig:purity_g_r_cut}.
+Compared to Figure~\ref{fig:purity_z} and \ref{fig:purity_g_r},
+the performance of BMC becomes worse in some photo-$z$ and $g-r$ bins.
+However, this drop in performance seems to be confined to
+only a small number of objects in regions of particular parameter space,
+and BMC still outperforms both TPC and morphological separation
+for the majority of objects.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/purity_z_cut.pdf}
+  \caption{Similar to Figure~\ref{fig:purity_mag_cut}
            but as a function of photo-$z$.}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/purity_g-r_cut.eps}
-  \caption{The same figure as Figure~\ref{purity_mag_cut},
+  \includegraphics[width=\columnwidth]{figures/purity_g_r_cut.pdf}
+  \caption{Similar to Figure~\ref{fig:purity_mag_cut}
            but as a function of $g-r$ color.}
-  \label{fig:purity_g-r_cut}
+  \label{fig:purity_g_r_cut}
 \end{figure}
 
-We present the star and galaxy purity values
-of Table~\ref{table:metrics_vvds} as a function of
-$i$-band magnitude in Figure~\ref{fig:purity_mag_cut}.
-It is clear that BMC outperforms both
-morphological separation and TPC classification techniques
-over all magnitudes.
-The galaxy purity of BMC is significantly better
-than that of TPC at bright magnitudes, $i\lesssim21.5$,
-while they are similar at faint magnitudes.
-The star purity of BMC shows improvement over that of TPC
-at faint magnitudes,
-while they are similar at bright magnitudes.
-As suggested by the weight maps in Figure~\ref{fig:weights_cut},
-BMC can accomplish this by combining information from
-all base classifiers,
-\eg giving more weight to the morphological separation method
-at bright magnitudes.
-We can also see in the bottom two panels that
-the performance of morphological separation suffers
-at intermediate magnitudes $21\lesssim i \lesssim23$,
-while BMC shows consistently better performance
-over all magnitude ranges.
-
-We also show the star and galaxy purity values
-of Table~\ref{table:metrics_vvds} as a function of
-photometric redshift in Figure~\ref{fig:purity_z}.
-Photo-$z$ is estimated with
-the \texttt{BPZ} algorithm~\citep{benitez2000bayesian}
-and provided with
-the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
-It is again clear that BMC outperforms
-both individual classification techniques over all redshift range.
-
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/p_dist_all.eps}
-    \caption{Histogram of the posterior probability that
-             a source is a star for morphological separation (top),
-             TPC (middle), and BMC (bottom) techniques
-             for the entire training data
-             (the CFHTLenS data in the DEEP2, SDSS, VIPERS, and VVDS
-             fields as discussed in Section~\ref{section:rich_training}).}
-    \label{fig:p_dist_all}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/p_dist_cut.eps}
-    \caption{The same figures as Figure~\ref{fig:p_dist_all},
-             but for TPC and BMC trained with 
-             poor training data
-             (the CFHTLenS data in the VDDS field
-             with a magnitude cut of $i<22$).}
+Compared to Figure~\ref{fig:p_dist_all},
+the difference between the posterior star probability distribution of
+TPC and that of BMC is now more pronounced in Figure~\ref{fig:p_dist_cut}.
+The $P\left(S\right)$ distribution of BMC for true galaxies
+falls off sharply at $P\left(S\right)\approx0.95$,
+and does not assign a star probability $P(S) \ga 0.95$ to any true galaxies,
+while both morphological separation and TPC techniques
+classify some true galaxies as stars with absolute certainty.
+
+\begin{figure}
+    \centering
+  \includegraphics[width=\columnwidth]{figures/p_dist_cut.pdf}
+  \caption{Similar to Figure~\ref{fig:p_dist_all},
+           but for the reduced training data set.}
   \label{fig:p_dist_cut}
-  \end{minipage}
-\end{figure*}
-
-We show the distribution of posterior star probabilities
-in Figures~\ref{fig:p_dist_all} and \ref{fig:p_dist_cut}.
-The histogram of spectroscopic galaxies is in blue,
-and the histogram of true stars is in red.
-It is interesting that the BMC technique assigns
-a probability $P\left(S\right)$ greater than 0.6
-to almost no true galaxies in Figure~\ref{fig:p_dist_all},
-presumably because BMC utilizes information from
-different types of classification techniques.
-This pattern is more pronounced in Figure~\ref{fig:p_dist_cut},
-where the $P\left(S\right)$ distribution of BMC for true galaxies
-falls off sharply at $P\left(S\right)\approx0.95$,
-while morphological separation and TPC techniques
-classify some true galaxies as stars with absolute certainty.
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\linewidth]{figures/color_color.eps}
-  \caption{Color-color diagrams of sources with
-           $0.45<P\left(S\right)<0.55$ when the BMC technique
-           is applied to the poor training data.}
-  \label{fig:color_color}
-\end{figure*}
-
-To identify the regions in color-color space
-where our BMC technique struggles to classify stars and galaxies,
-we present the distribution over the color-color space for sources with
-$0.45 < P(S) < 0.55$ in Figure~\ref{fig:color_color}.
-Not surprisingly, these regions correspond to
-where there is a significant overlap between stars and galaxies
-in the color space.
-It may be possible to improve the performance of
-star-galaxy classification by obtaining more information
-through spectroscopic follow-up, manual study, or citizen science projects,
-or by developing a classification technique specifically designed
-for these regions of the color space and
-combining it with other techniques in a Bayesian framework.
+\end{figure}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=\columnwidth]{figures/color_color.eps}
+%  \caption{Color-color diagrams of sources with
+%           $0.45<P\left(S\right)<0.55$ when the BMC technique
+%           is applied to the poor training data.}
+%  \label{fig:color_color}
+%\end{figure*}
+%
+%To identify the regions in color-color space
+%where our BMC technique struggles to classify stars and galaxies,
+%we present the distribution over the color-color space for sources with
+%$0.45 < P(S) < 0.55$ in Figure~\ref{fig:color_color}.
+%Not surprisingly, these regions correspond to
+%where there is a significant overlap between stars and galaxies
+%in the color space.
+%It may be possible to improve the performance of
+%star-galaxy classification by obtaining more information
+%through spectroscopic follow-up, manual study, or citizen science projects,
+%or by developing a classification technique specifically designed
+%for these regions of the color space and
+%combining it with other techniques in a Bayesian framework.
 
@@
 
-It is very costly in terms of telescope time to
-obtain a large sample spectroscopic observations
-down to the limiting magnitude of a photometric sample.
-Thus, we investigate the impact of training set quality
-by imposing various magnitude and $z$ cuts
-to decrease the size of the training set.
-
-In Figure.~\ref{fig:perform_mag_cut}, we present
-a visual comparison between different classification techniques
-when various magnitude cuts are applied.
-
-In Figure.~\ref{fig:perform_z_cut}, we also present
-the result of imposing various $z$ cuts.
-
-\begin{figure*}
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
-    \caption{The classification performance metrics for TPC, HB, morphology,
-      and BMC as applied to the CFHTLenS data in the VVDS field
+The combination techniques that we have demonstrated in this work uses
+two training based algorithms as base classifiers.
+Ideally, the training data should mirror the entire parameter space
+occupied by the data to be classified.
+Yet we have seen in Section~\ref{section:poor_training}
+that the BMC technique does reliably
+extrapolate past the limits of the training data,
+even when some base classifiers are trained on a low-quality training data set.
+In this section, we further investigate
+if and where BMC begins to break down
+by imposing various magnitude, photo-$z$, and color cuts
+to change the size and composition of the training set.
+
+In Figure~\ref{fig:perform_mag_cut}, we present
+a visual comparison between different classification techniques,
+when various magnitude cuts are applied in the training data,
+and the performance is measured on the same test set
+from Section~\ref{section:rich_training} and \ref{section:poor_training}.
+It is not surprising that the performance of TPC decreases
+as we decrease the size of training set
+by imposing more restrictive magnitude cuts,
+while the performance of HB and morphological separation
+is essentially unchanged.
+However, the effect of change in size and composition of the training set
+is significantly mitigated by the use of the BMC technique.
+BMC outperforms both HB and TPC in all four metrics,
+even when training is restricted to $i < 20.0$.
+BMC also consistently outperforms morphological separation
+until we impose a magnitude cut of $i < 20.0$ on the training data,
+beyond which point BMC finally performs worse than morphological separation.
+It is remarkable that BMC is able to reliably extrapolate
+past the training data to $i \sim 24.5$,
+the limiting magnitude of the test set, and outperform HB, TPC,
+and morphological separation in all performance metrics,
+even the demographics of training set do not accurately sample 
+the data to be classified.
+
+Similarly, we impose various spectroscopic redshift cuts 
+on the training data in Figure~\ref{fig:perform_z_cut}. 
+BMC begins to perform worse than morphological separation
+when a conservative cut of $z_{\rmn{spec}} < 0.6$ is imposed.
+However, it is again clear that BMC is able to
+utilize information from HB and morphological separation to
+mitigate the drop in the performance of TPC.
+
+In Figure~\ref{fig:perform_g_r_cut},
+we decrease the size of training set
+by keeping red objects and gradually removing blue objects.
+A color cut seems to have a more pronounced effect
+on the performance of TPC and BMC,
+and BMC performs worse than morphological separation
+when the training set is restricted to $g - r > 0.4$.
+The performance depends more strongly on the color distribution, because 
+a significant fraction of blue objects consists of stars,
+while objects with fainter magnitudes and higher redshifts
+are mostly galaxies.
+We can verify this in Figure~\ref{fig:som_colors},
+where the darker (higher stellar fraction) cells in the upper middle region
+of the stellar fraction map (top right)
+have bright magnitudes $i \la 20$ in the $i$-band magnitude map (top left)
+and blue colors $g-r \la 0.5$ in the $g-r$ color map (bottom right).
+On the other hand, the darker (fainter magnitude) cells
+in the lower left corner of the $i$-band magnitude map
+have almost no stars in them and
+are represented by bright (low stellar fraction)
+cells in the stellar fraction map.
+Thus, these results indicate that the performance of training based
+methods depends more strongly on the composition of training data
+than on the size, and it is necessary to have
+a sufficient number of the minority class in the training data set
+to ensure optimal performance.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/perform_mag_cut.pdf}
+    \caption{The classification performance metrics for
+      BMC (blue), TPC (green), morphology (red), and HB (purple)
+      as applied to the CFHTLenS data in the VVDS field
       with various magnitude cuts.
@@
     \label{fig:perform_mag_cut}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[t]{0.49\linewidth}
-    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
-    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
-      $z_{\text{spec}}$ cuts.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/perform_z_cut.pdf}
+    \caption{Similar to Figure~\ref{fig:perform_mag_cut}
+        but using $z_{\text{spec}}$ cuts.}
     \label{fig:perform_z_cut}
-  \end{minipage}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/perform_g_r_cut.pdf}
+    \caption{Similar to Figure~\ref{fig:perform_mag_cut}
+        but using $g - r$ color cuts.}
+    \label{fig:perform_g_r_cut}
+\end{figure}
+
 \section{Conclusions}
@@
 
-\textcolor{blue}{We
-have presented and analyzed a novel star-galaxy classification framework
-for combining classifications from the CFHTLenS project.
-In particular, we use four independent star-galaxy classification techniques:
+We have presented and analyzed a novel star-galaxy classification framework
+for combining star-galaxy classifiers using the CFHTLenS data.
+In particular, we use four independent classification techniques:
 a morphological separation method;
-\texttt{TPC}, a supervised machine learning technique
+TPC, a supervised machine learning technique
 based on prediction trees and a random forest;
-\texttt{SOMc}, an unsupervised machine leraning approach
+SOMc, an unsupervised machine learning approach
 based on self-organizing maps and a random atlas;
-and \texttt{HB}, a hierarichical bayesian template-fitting method
-that we have modified and parallized.
-Both \texttt{TPC} and \texttt{SOMc} are currently availble within
+and HB, a Hierarchical Bayesian template-fitting method
+that we have modified and parallelized.
+Both TPC and SOMc algorithms are currently available within
 a software package named
-\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}.
-Our implementaion of \texttt{HB} and \texttt{BMC},
-as well as IPython notebooks that were used to
+\textsc{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}.
+Our implementation of HB and BMC,
+as well as \textsc{IPython} notebooks that have been used to
 produce the results in this paper,
-are available at \url{https://github.com/EdwardJKim/astroclass}.  }
-
-When there's wealth of training data,
-TPC is so good that combination techniques show
-no significant improvement over TPC.
-
-Even with poor (but reasonable) training data,
-TPC outperforms unsupervised methods.
-
-In a more realistic case where the training data is poor,
-BMC should be used.
-
-Probabilistic classification is better.
-Binary classification can be transformed into a probabilistic
-classification when combined with other models in a Bayesian framework.
-
-We note that \cite{carrascokind2014exhausting} analyzed
+are available at \url{https://github.com/EdwardJKim/astroclass}.
+
+Given the variety of star-galaxy classification methods we are using,
+we fully expect the relative performance
+of the individual techniques to vary across
+the parameter space spanned by the data.
+We therefore adopt the binning strategy, where
+we allow different classifier combinations 
+in different parts of parameter space
+by creating two-dimensional self-organizing maps of
+the full multi-dimensional magnitude-color space.
+We apply different star-galaxy classification techniques
+within each cell of this map,
+and find that the four techniques are weighted most strongly
+in different regions of the map.
+
+Using data from the CFHTLenS survey,
+we have considered different scenarios:
+when an excellent training set is available with spectroscopic labels from
+DEEP2, SDSS, VIPERS, and VVDS, and
+when the demographics of sources in a low-quality training set
+do not match the demographics of objects in the test data set.
+We demonstrate that the Bayesian Model Combination (BMC) technique improves
+the overall performance over any individual classification method
+in both cases.
+We note that \citet{carrascokind2014exhausting} analyzed
 different techniques for combining
-photometric redshift probability density functions (photo-$z$ PDFs),
+photometric redshift probability density functions (photo-$z$ PDFs)
 and also found that BMC is in general the best
@@
 
-\textcolor{red}{more here}
+The combination technique described in this paper
+can be easily applied to other surveys.
+Given the efficacy of our approach,
+classifier combination strategies are likely the optimal approach
+for currently ongoing and forthcoming photometric surveys.
+We therefore plan to apply the combination technique described in this paper
+to other surveys such as the DES.
+Our approach can also be extended more broadly to
+classify objects that are neither stars nor galaxies (\eg quasars).
+Finally, future studies could explore the use of multi-epoch data,
+which would be particularly useful for the next generation of
+synoptic surveys.
 
@@
 
-We gratefully acknowledge the use of computing resources
-from the Extreme Science and Engineering Discovery Environment (XSEDE),
-which is supported by National Science Foundation grant number
-\textcolor{red}{XXXX}.
-
-Funding for CFHTLenS has been provided by
-\textcolor{red}{XXXX}.
-
+This work used the Extreme Science and Engineering Discovery Environment
+(XSEDE), which is supported by National Science Foundation grant number
+ACI-1053575.
+
+This work is based on observations obtained with MegaPrime/MegaCam, a
+joint project of CFHT and CEA/DAPNIA, at the Canada-France-Hawaii
+Telescope (CFHT) which is operated by the National Research Council
+(NRC) of Canada, the Institut National des Sciences de l'Univers of
+the Centre National de la Recherche Scientifique (CNRS) of France, and
+the University of Hawaii. This research used the facilities of the
+Canadian Astronomy Data Centre operated by the National Research
+Council of Canada with the support of the Canadian Space Agency.
+CFHTLenS data processing was made possible thanks to significant
+computing support from the NSERC Research Tools and Instruments grant
+program.
+
+Funding for the DEEP2 survey has been provided by NSF grants AST-0071048,
+AST-0071198, AST-0507428, and AST-0507483.
+
+Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the
+Participating Institutions, the National Science Foundation, and the U.S.
+Department of Energy Office of Science. The SDSS-III web site is
+http://www.sdss3.org/.
+
+SDSS-III is managed by the Astrophysical Research Consortium for the
+Participating Institutions of the SDSS-III Collaboration including the
+University of Arizona, the Brazilian Participation Group, Brookhaven National
+Laboratory, Carnegie Mellon University, University of Florida, the French
+Participation Group, the German Participation Group, Harvard University, the
+Instituto de Astrofisica de Canarias, the Michigan State/Notre Dame/JINA
+Participation Group, Johns Hopkins University, Lawrence Berkeley National
+Laboratory, Max Planck Institute for Astrophysics, Max Planck Institute for
+Extraterrestrial Physics, New Mexico State University, New York University,
+Ohio State University, Pennsylvania State University, University of Portsmouth,
+Princeton University, the Spanish Participation Group, University of Tokyo,
+University of Utah, Vanderbilt University, University of Virginia, University
+of Washington, and Yale University.
+
+This paper uses data from the 
+VIMOS Public Extragalactic Redshift Survey (VIPERS).
+VIPERS has been performed using the ESO Very Large Telescope, under the "Large
+Programme" 182.A-0886. The participating institutions and funding agencies are
+listed at http://vipers.inaf.it/.
+
+This research uses data from the VIMOS VLT Deep Survey, obtained from the VVDS
+database operated by Cesam, Laboratoire d'Astrophysique de Marseille, France.
+
+\footnotesize{
 \bibliographystyle{mn2e}
-
 \bibliography{sg_paper}
+}
 

### 2015-05-08 paper/main.tex
udpate paper

@@
 an unsupervised machine learning method based on self-organizing maps,
-and a Hierarchical Bayesian template fitting method.
+and a hierarchical Bayesian template fitting method.
 %Any of these four methods could
@@
 In Section~\ref{section:classification_combination_methods},
-we describe different classification combination techniques
+we describe different classification combination techniques.
 In Section~\ref{section:data},
@@
 The fourth method is
-a Hierarchical Bayesian template fitting technique
+a Hierarchical Bayesian (HB) template fitting technique
 based on the work by \citet{Fadely2012},
@@
   \centering
-  \includegraphics[width=\columnwidth]{figures/morph.eps}
+  \includegraphics[width=\columnwidth]{figures/morph.pdf}
   \caption{Half-light radius vs.\ magnitude.}
@@
 the input photometry
-and determines a probability estimate by computing
+and determine a probability estimate by computing
 the posterior that a source is a star or a galaxy.
@@
 Similarly, we also parametrize the overall prior probability,
-$\left(S | \bmath{\theta}\right)$ in Equation~\ref{eq:overall_posterior},
+$\left(S | \bmath{\theta}\right)$, in Equation~\ref{eq:overall_posterior},
 as a weight.
@@
 individual star-galaxy classification techniques
-using four combination techniques,
-including Bayesian Model Combination (BMC)
-and a variant of an ensemble learning technique known as stacking.
+using four combination techniques.
 The main idea behind ensemble learning is to weight
@@
 such as SOM or template fitting approaches 
-to generally perform better when training data
+to generally perform better when a training sample
 is either sparse or unavailable.
@@
 to compute the relative weights for the base classifiers.
-The weights within individual cells are then applied to the test data
-to make the prediction.
+The weights within individual cells are then applied to
+the blind test data set to make the prediction.
 
@@
 and $y_{i}$ is the probability prediction made by the models.
-Thus, a model with the minimum MSE is chosen in each bin
+Thus, a model with the minimum MSE is chosen in each bin,
 and is assigned a weight of one,
@@
   P \left( e | \bmath{D} \right) \propto 
-  P \left( e \right) \left(1 - \epsilon \right)^{N - M_s + M_g}
+  P \left( e \right) \left(1 - \epsilon \right)^{N - M_s - M_g}
   \left( \epsilon \right)^{M_s + M_g}.
@@
 It covers a total of 154 square degrees
-in the four fields of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
+in the four fields (named W1, W2, W3, and W4)
+of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
 observed in the five photometric bands:
@@
 For example, \cite{hildebrandt2012cfhtlens} suggest that
-all objects with $i > 23$ in the CFHTLenS dataset
+all objects with $i > 23$ in the CFHTLenS data set
 may be classified as galaxies
@@
 will meet the science requirements for the quality of
-star-galaxy classification in lower latitude, star-crowded fields.
-Rather, our goal for the CFHTLenS dataset
+star-galaxy classification in lower-latitude, star-crowded fields.
+Rather, our goal for the CFHTLenS data set
 is to demonstrate the usefulness
@@
 \noindent
-where $N_g$ is the number of true galaxies classified as galaxies
+where $N_g$ is the number of true galaxies classified as galaxies,
 and $M_g$ is the number of true galaxies classified as stars.
@@
 which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
-at a certain purity ($p_g = 0.9484$, $p_s = 0.9666$),
+at a certain purity ($p_g = 0.9597$, $p_s = 0.9666$),
 we adjust the threshold of probabilistic classifiers
@@
 
-In Figure~\ref{fig:som_colors}, we show
-the mean CFHTLenS $i$-band magnitude in the top left panel,
-and the fraction of stars in each cell in the top right panel.
+In Figure~\ref{fig:som_colors}, we show in the top left panel
+the mean CFHTLenS $i$-band magnitude in each cell,
+and in the top right panel the fraction of stars in each cell.
 The bottom two panels show the mean $u-g$ and  $g-r$ colors in each cell.
@@
 at bright magnitudes.
-The bottom panels shows that
+The bottom panel shows that
 the star purity of morphological separation drops to $p_s < 0.8$
@@
 $i > 21$.
-BMC clearly outperforms TPC, the benchmark base classifier,
+BMC clearly outperforms both TPC and morphological separation,
 and it maintains the overall galaxy purity of 0.980
@@
 
-We also show the star and galaxy purity values as a function of
+We also show the star and galaxy purity values as functions of
 photometric redshift estimate in Figure~\ref{fig:purity_z}.
@@
 it is clear that BMC outperforms
-both TPC and morphological separation over all redshift range.
+both TPC and morphological separation over all redshifts.
 We also present in Figure~\ref{fig:purity_g_r}
@@
 It is again clear that BMC outperforms
-both TPC and morphological separation over all $g-r$ color range.
+both TPC and morphological separation over all $g-r$ colors.
 
@@
            and the four different classification combination methods
-           for when the training data set consists of
+           when the training data set consists of
            only the sources that are in CFHTLS W1 field,
@@
 Alhtough BoM always chooses TPC as the best model
-when there is a high-quality training set,
-it now chooses various methods and outperforms all base classifiers.
-While the performance of the stacking technique is comparable
-to that of BMC in Section~\ref{section:rich_training},
+when we have a high-quality training set,
+it now chooses various methods in different bins
+and outperforms all base classifiers.
+While the performance of the stacking technique is only slightly worse
+than that of BMC when we have a high-quality training set,
 stacking now fails to outperform morphological separation.
@@
 that of the test set are different.
-Since the training set is cut at $i < 22$
+Since the training set is cut at $i < 22$,
 the density distribution falls off sharply around $i \sim 22$
 and has a higher fraction of stars than the test set.
-Compared to Figure~\ref{fig:purity_mag},
-the purity of TPC suffers a significant decrease in purity.
+Compared to the purity values in Figure~\ref{fig:purity_mag},
+TPC now suffers a significant decrease in star and galaxy purity.
 However, the purity of BMC does not show such a significant drop
@@
 However, this drop in performance seems to be confined to
-only a small number of objects in regions of particular parameter space,
+only a small number of objects in particular regions of
+the parameter space,
 and BMC still outperforms both TPC and morphological separation
@@
 falls off sharply at $P\left(S\right)\approx0.95$,
-and does not assign a star probability $P(S) \ga 0.95$ to any true galaxies,
-while both morphological separation and TPC techniques
+and BMC does not assign a star probability
+$P(S) \ga 0.95$ to any true galaxies,
+On the other hand, both TPC and morphological separation 
 classify some true galaxies as stars with absolute certainty.
@@
   \includegraphics[width=\columnwidth]{figures/p_dist_cut.pdf}
-  \caption{Similar to Figure~\ref{fig:p_dist_all},
+  \caption{Similar to Figure~\ref{fig:p_dist_all}
            but for the reduced training data set.}
@@
 
-The combination techniques that we have demonstrated in this work uses
+The combination techniques that we have demonstrated so far use
 two training based algorithms as base classifiers.
@@
 a visual comparison between different classification techniques,
-when various magnitude cuts are applied in the training data,
+when various magnitude cuts are applied on the training data,
 and the performance is measured on the same test set
@@
 BMC outperforms both HB and TPC in all four metrics,
-even when training is restricted to $i < 20.0$.
+even when the training set is restricted to $i < 20.0$.
 BMC also consistently outperforms morphological separation
@@
 on the performance of TPC and BMC,
-and BMC performs worse than morphological separation
+which perform worse than morphological separation
 when the training set is restricted to $g - r > 0.4$.
@@
 where the darker (higher stellar fraction) cells in the upper middle region
-of the stellar fraction map (top right)
-have bright magnitudes $i \la 20$ in the $i$-band magnitude map (top left)
-and blue colors $g-r \la 0.5$ in the $g-r$ color map (bottom right).
+of the stellar fraction map (top right panel)
+have bright magnitudes $i \la 20$
+in the $i$-band magnitude map (top left panel)
+and blue colors $g-r \la 0.5$ in the $g-r$ color map (bottom right panel).
 On the other hand, the darker (fainter magnitude) cells
-in the lower left corner of the $i$-band magnitude map
+in the right-hand side of the $i$-band magnitude map
 have almost no stars in them and
@@
 than on the size, and it is necessary to have
-a sufficient number of the minority class in the training data set
-to ensure optimal performance.
+a sufficient number of the minority class
+in the training data set to ensure optimal performance.
 

### 2015-05-08 paper/main.tex
minor typo - third person singular of have = has

@@
            only the sources that are in CFHTLS W1 field,
-           have spectroscopic labels available from VVDS,
-           and have $i < 22$.
+           has spectroscopic labels available from VVDS,
+           and has $i < 22$.
            The definition of the metrics is summarized in

### 2015-05-08 paper/main.tex
update paper

@@
 until the galaxy completeness $c_g$ matches
-that of morphological separtion
+that of morphological separation
 to compute the galaxy purity $p_{g}$ at $c_g=0.9964$.
@@
            blue, green, and red, respectively.
-           The $1 \sigma$ confidence bands are estaimted by
+           The $1 \sigma$ confidence bands are estimated by
            bootstrap sampling.}
@@
 %all individual classification techniques,
-Alhtough BoM always chooses TPC as the best model
+Although BoM always chooses TPC as the best model
 when we have a high-quality training set,
@@
 the drop in the performance of TPC is clear.
-However, even when some classifers have been trained on
+However, even when some classifiers have been trained on
 a significantly reduced training set,
@@
 
-We thank the referee for a careful reading of the manuscript
-and comments that improved this work.
 We thank Ignacio Sevilla for helpful and insightful conversations.
-We acknowledge support from the National Science Grant No.
-\textcolor{red}{XXXX}.
+We acknowledge support from the 
+National Science Foundation Grant No.\ AST-1313415.
+RJB acknowledges support as an Associate
+within the Center for Advanced Study at the University of Illinois.
 

### 2015-05-08 paper/main.tex
add stats affiliation to RJB

@@
 \author[E. J. Kim, R. J. Brunner \& M. Carrasco Kind]{
-  Edward J.~Kim$^1$\thanks{jkim575@illinois.edu}, Robert J.~Brunner$^2$,
-  and Matias Carrasco Kind$^{2,3}$\\
+  Edward J.~Kim$^1$\thanks{jkim575@illinois.edu},
+  Robert J.~Brunner$^{2,3,4}$,
+  and Matias Carrasco Kind$^{2,4}$\\
 $^1$Department of Physics, University of Illinois, Urbana, IL 61801 USA\\
 $^2$Department of Astronomy, University of Illinois, Urbana, IL 61801 USA\\
-$^3$National Center for Supercomputing Applications, Urbana, IL 61801 USA}
+$^3$Department of Statistics, University of Illinois, Champaign, IL 61820 USA\\
+$^4$National Center for Supercomputing Applications, Urbana, IL 61801 USA}
 

### 2015-05-11 bmc/bmc.py
commit BMC source code

@@ -1,0 +1,207 @@
+'''
+Author: Edward J. Kim
+Email: jkim575@ilinois.edu
+'''
+from __future__ import print_function, division, unicode_literals
+import numpy as np
+import os
+from mlz.ml_codes import SOMZ
+
+class BMC(object):
+    '''
+    Bayesian Model Combination (Monteith et al. 2011).
+    A Hybrid Ensemble Approach to Star-galaxy Classification (Kim, Brunner & Carrasco Kind 2015).
+    '''
+    def __init__(self, q=3, ncomb=1200, nburn=200, ntop=10, icore=0):
+        '''
+        Constructor.
+        
+        Parameters
+        ----------
+        q: The number of combinations drawn each time from Dirichlet distribution.
+        ncomb: The total number of ensembles.
+        nburn: The number of iterations in the burn-in step.
+        ntop: The side length of a rectangular SOM.
+        icore: The number of cores (0 unless parallelized).
+        '''
+        self.q = q
+        self.ncomb = ncomb
+        self.nburn = nburn
+        self.ntop = ntop
+        self.icore = icore
+        self.pfloor = 1e-300
+
+    def get_SOM_cells(self, topology='grid', ntop=10, iterations=200,
+                      periodic='no', iproc=0):
+        '''
+        Creates a SOM representation.
+        Only supports rectangular topology.
+        '''
+        if iproc is not 0:
+            print('Creating {}th SOM map...'.format(iproc))
+
+        if topology == 'grid':
+            self.n_cell = ntop**2
+
+        # Calls the SOMZ mode
+        self.M = SOMZ.SelfMap(self.X_train, self.Y_train,
+                         topology=topology, Ntop=ntop,
+                         iterations=iterations, periodic=periodic)
+        # creates a map
+        self.M.create_mapF()
+        # evaluates it with the Y entered, or anyoher desired colum
+        self.M.evaluate_map()
+
+        cells_train = np.zeros(len(self.X_train))
+
+        for i in range(len(self.X_train)):
+            cells_train[i] = self.M.som_best_cell(self.X_train[i])[0]
+
+        np.save(self.file_train_cells, cells_train)
+
+    def get_log_likelihood(self, label, array, weight):
+        '''
+        The likelihood function.
+        
+        Parameters
+        ----------
+        label: A numpy array. The ground truth.
+        array: The (probabilistic) outputs of base classifiers.  
+        weight: The weights of base classifiers.
+        
+        Returns
+        -------
+        A numpy array.
+        '''
+        y = label
[…]

### 2015-05-11 docker/notebook.sh
add Dockerfile

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+# Strict mode
+set -euo pipefail
+IFS=$'\n\t' 
+
+# Create a self signed certificate for the user if one doesn't exist
+if [ ! -f $PEM_FILE ]; then
+  openssl req -x509 -nodes -days 365 -newkey rsa:2048 -keyout $PEM_FILE -out $PEM_FILE \
+    -subj "/C=XX/ST=XX/L=XX/O=dockergenerated/CN=dockergenerated"
+fi
+
+# Create the hash to pass to the IPython notebook, but don't export it so it doesn't appear
+# as an environment variable within IPython kernels themselves
+HASH=$(python3 -c "from IPython.lib import passwd; print(passwd('${PASSWORD}'))")
+unset PASSWORD
+
+CERTFILE_OPTION="--certfile=$PEM_FILE"
+
+if [ $USE_HTTP -ne 0 ]; then
+  CERTFILE_OPTION=""
+fi
+
+ipython2 notebook --no-browser --port 8889 --ip=* $CERTFILE_OPTION --NotebookApp.password="$HASH" --matplotlib=inline

### 2015-05-11 hbsgc/calc_utils.py
add hbsgc files

@@ -1,0 +1,22 @@
+'''
+Matrix calculation utils for HBSGC.
+'''
+
+import numpy as np
+
+
+def dot(left_array, right_array, ncols=1):
+    '''
+    Performs matrix dot product column by column to avoid memory error.
+    '''
+    m, left_inner = left_array.shape
+    right_inner, n = right_array.shape
+
+    if not left_inner == right_inner:
+        raise ValueError('shapes are not alighed')
+
+    out = np.zeros((m, n))
+    for i in xrange(0, n, ncols):
+        out[:, i] = np.dot(left_array, right_array[:, i])
+
+    return out

### 2015-07-09 data/script/download_all.py
add download script

@@ -1,0 +1,54 @@
+#!/usr/bin/env python
+from __future__ import print_function
+import urllib2
+
+def main():
+    '''
+    A quick script that dumps all columns of all the WIDE fields in the CFHTLenS survey
+    to the current directory
+    '''
+
+    field_list = ['W1m0m0', 'W1m0m1', 'W1m0m2', 'W1m0m3', 'W1m0m4', 'W1m0p1',
+                  'W1m0p2', 'W1m0p3', 'W1m1m0', 'W1m1m1', 'W1m1m2', 'W1m1m3',
+                  'W1m1m4', 'W1m1p1', 'W1m1p2', 'W1m1p3', 'W1m2m0', 'W1m2m1',
+                  'W1m2m2', 'W1m2m3', 'W1m2m4', 'W1m2p1', 'W1m2p2', 'W1m2p3',
+                  'W1m3m0', 'W1m3m1', 'W1m3m2', 'W1m3m3', 'W1m3m4', 'W1m3p1',
+                  'W1m3p2', 'W1m3p3', 'W1m4m0', 'W1m4m1', 'W1m4m2', 'W1m4m3',
+                  'W1m4m4', 'W1m4p1', 'W1m4p2', 'W1m4p3', 'W1p1m0', 'W1p1m1',
+                  'W1p1m2', 'W1p1m3', 'W1p1m4', 'W1p1p1', 'W1p1p2', 'W1p1p3',
+                  'W1p2m0', 'W1p2m1', 'W1p2m2', 'W1p2m3', 'W1p2m4', 'W1p2p1',
+                  'W1p2p2', 'W1p2p3', 'W1p3m0', 'W1p3m1', 'W1p3m2', 'W1p3m3',
+                  'W1p3m4', 'W1p3p1', 'W1p3p2', 'W1p3p3', 'W1p4m0', 'W1p4m1',
+                  'W1p4m2', 'W1p4m3', 'W1p4m4', 'W1p4p1', 'W1p4p2', 'W1p4p3',
+                  'W2m0m0', 'W2m0m1', 'W2m0p1', 'W2m0p2', 'W2m0p3', 'W2m1m0',
+                  'W2m1m1', 'W2m1p1', 'W2m1p2', 'W2m1p3', 'W2p1m0', 'W2p1m1',
+                  'W2p1p1', 'W2p1p2', 'W2p1p3', 'W2p2m0', 'W2p2m1', 'W2p2p1',
[…]

### 2015-07-10 paper/main.tex
the submitted version of the paper
update paper

@@
 
-\title{A Hybrid Ensemble Learning Approach to Star-Galaxy Classification}
+\title[A Hybrid Approach to Star-Galaxy Classification]{A Hybrid Ensemble Learning Approach to Star-Galaxy Classification}
 
@@
 \begin{abstract}
-%Star-galaxy separation in currently ongoing and future
-%large photometric surveys, such as DES and LSST,
-%will detect a vast number of unresolved galaxies at faint magnitudes.
 There exist a variety of star-galaxy classification techniques,
@@
 and a hierarchical Bayesian template fitting method.
-%Any of these four methods could
-%easily be replaced by a similar method
-%or be extended to include other existing techniques.
 Using data from the CFHTLenS survey,
@@
 to tackle this long-standing problem,
-%and yet it remains an unsolved challenge.
 and yet there is no consensus on
@@
 various properties of the input data.
-The differences between a SOM and other neural network algorithms are
+The differences between a SOM and
+typical neural network algorithms are
 that a SOM is unsupervised,
@@
 the training data is separated into groups of similar features.
+Although the spectroscopic labels are not used at all
+in the learning process, they are used
+(only after the map has been constructed)
+to generate predictions for each cell in the resulting two-dimensional map.
 
@@
 which can be rectangular, hexagonal, or spherical.
-To classify stars and galaxies in the CFHTLenS data,
-we use a spherical topology,
-which is constructed by using
+In our SOM implementation, it is also possible
+to use periodic boundary conditions for the non-spherical cases.
+The spherical topology is by definition periodic and
+is constructed by using
 \textsc{HEALPIX}~\citep{gorski2005healpix}.
-Furthermore, similar to TPC,
+Similar to TPC,
 we use the OOB technique to make an unbiased estimation of errors.
+We determine the optimal parameters by performing
+a grid search in the parameter space of different toplogies,
+as well as other SOM parameters, for the OOB data.
+We find that the spherical topology gives the best performance
+for the CFHTLenS data, likely due to its natural periodicity.
+Thus, we use a spherical topology to classify stars and galaxies
+in the CFHTLenS data.
 For a complete description of the SOM implementation
@@
 visualization as shown in Figure~\ref{fig:som_colors}.
+We note that this choice is mainly for convenience
+and that the optimal topology and map size would likely depend on
+a number of factors, such as the number of objects and attributes.
 For all combination methods,
@@
 the VIMOS Public Extragalactic Redshift
-Survey~\citep[VIPERS;][]{garilli2014vimos},
+Survey~\citep[VIPERS;][]{garilli2014vimos}.
+We have selected only sources with very secure
+redshifts and no bad flags (quality flags -1, 3, and 4 for DEEP2;
+quality flag 0 for SDSS; quality flags 3, 4, 23, and 24 for VIPERS
+and VVDS).
 In the end, we have 8,545 stars and 57,843 galaxies available
@@
 and use the remainder for training and cross-validation.
+While HB uses only the magnitudes in the five bands,
+$u$, $g$, $r$, $i$, and $z$,
+TPC and SOMc are trained with a total of 9 attributes: the five magnitudes
+and their corresponding colors, $u-g$, $g-r$, $r-i$, and $i-z$.
+The morphological separation method uses \textsc{SExtractor}'s 
+FLUX\_RADIUS parameter provided by the CFHTLenS catalog.
 
@@
 the base classifiers may be poor or trained on partial data.
+We also note that the relatively few number of stars
+in the CFHTLS fields might paint too positive a picture of completeness and purity,
+especially for the stars.
+Thus, we caution the reader that the specific completeness and purity values
+will likely vary in other surveys that observe large portions of the sky,
+and we emphasize once again that our aim is to highlight that there is a
+relative improvement in performance when we combine multiple star-galaxy
+classification techniques to generate a meta-classification.
 
@@
            The bold entries highlight the best performance values
-           within each column.}
+           within each column.
+           Note that some objects in the test set have bad or missing
+           values (\eg $-99$ or $99$) in one or more attributes,
+           which are included here (but are omitted, for example,
+           in Figure~\ref{fig:purity_mag_integrated} when the corrsponding
+           attribute is not available.)}
   \centering
@@
 for BMC, TPC, and morphological separation
-as functions of $i$-band magnitude for the differential counts.
+as functions of $i$-band magnitude.
 We use the kernel density estimation \cite[KDE;][]{silverman1986density}
@@
 
-In Figure~\ref{fig:purity_mag_integrated}, we show
-the overall galaxy and star purity values as functions of magnitude
-for the integrated counts.
+In Figure~\ref{fig:purity_mag_integrated}, we show the
+cumulative galaxy and star purity values as functions of magnitude.
 Although morphological separation performs better than TPC
@@
   \caption{Purity as a function of the $i$-band magnitude
-           as estimated by the kernel density estimation (KDE) method
-           for the differential counts.
+           as estimated by the kernel density estimation (KDE) method.
            The top panel shows the histogram with a bin size of 0.1 mag
@@
   \includegraphics[width=\columnwidth]{figures/purity_mag_integrated.pdf}
-  \caption{Purity as a function of the $i$-band magnitude
-           for the integrated counts.
+  \caption{Cumulative purity as a function of the $i$-band magnitude.
            The upper panel compares
@@
            The bold entries highlight the best performance values
-           within each column.}
+           within each column.
+           Note that some objects in the test set have bad or missing
+           values (\eg $-99$ or $99$) in one or more attributes,
+           which are included here (but are omitted, for example,
+           in Figure~\ref{fig:purity_mag_cut_integrated} when the corrsponding
+           attribute is not available.)}
   \centering
@@
 AUC, $p_s$ at $c_s=0.7145$, $p_g$ at $c_g=0.9600$, and $p_s$ at $c_s=0.2500$.
-%Interestingly, although TPC does not perform better
-%than morphological separation,
-%BoM still chooses TPC as the best model within each SOM cell
-%and shows exactly the same performance metrics as TPC.
-%Stacking and BMC techniques still outperform
-%all individual classification techniques,
 Although BoM always chooses TPC as the best model
@@
 
-%% ROC curve for vvds i < 20
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=\columnwidth]{figures/roc_cut.eps}
-%  \caption{A magnified view of the ROC curve of
-%           four star-galaxy individual classification approaches
-%           (TPC, SOMc, HB, and morphology)
-%           and the best-performing combination technique (BMC).
-%           trained with the reduced data set.
-%           We show only one of the four combination methods, BMC,
-%           which has the best overall performance.}
-%  \label{fig:roc_cut}
-%\end{figure}
-
-%In Figure~\ref{fig:roc_cut}, we plot the ROC curve
-%of all four individual classification algorithm
-%and our BMC technique for the poor training data case.
-%We note that a binary classifier,
-%such as our cut-based morphological separation,
-%whose output is either zero or one,
-%is a single point in the ROC space.
-%We vary the half-light radius to create a smooth ROC curve
-%for morphological separation.
-%It is clear that our BMC technique dominates
-%each individual classification techniques in the ROC space.
-
-%% weights map for vvds i < 20
-
 \begin{figure}
@@
 two-dimensional weight map defined by the SOM.
-%It seems that TPC performs well
-%in most of the parameter space.
 When the quality of training data is relatively poor,
@@
 We present the star and galaxy purity values as functions of
-$i$-band magnitude for the differential counts
-in Figure~\ref{fig:purity_mag_cut}.
+$i$-band magnitude in Figure~\ref{fig:purity_mag_cut}.
 The normalized density distribution as a function of magnitude
@@
 
-In Figure~\ref{fig:purity_mag_cut_integrated}, we show
-the overall galaxy and star purity values as functions of magnitude
-for the integrated counts.
+In Figure~\ref{fig:purity_mag_cut_integrated}, we show the
+cumulative galaxy and star purity values as functions of magnitude.
 Compared to Figure~\ref{fig:purity_mag_integrated},
@@
 \end{figure}
-
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=\columnwidth]{figures/color_color.eps}
-%  \caption{Color-color diagrams of sources with
-%           $0.45<P\left(S\right)<0.55$ when the BMC technique
-%           is applied to the poor training data.}
-%  \label{fig:color_color}
-%\end{figure*}
-%
-%To identify the regions in color-color space
-%where our BMC technique struggles to classify stars and galaxies,
-%we present the distribution over the color-color space for sources with
-%$0.45 < P(S) < 0.55$ in Figure~\ref{fig:color_color}.
-%Not surprisingly, these regions correspond to
-%where there is a significant overlap between stars and galaxies
-%in the color space.
-%It may be possible to improve the performance of
-%star-galaxy classification by obtaining more information
-%through spectroscopic follow-up, manual study, or citizen science projects,
-%or by developing a classification technique specifically designed
-%for these regions of the color space and
-%combining it with other techniques in a Bayesian framework.
 
@@
 on the training data in Figure~\ref{fig:perform_z_cut}.
+Since all stars have $z_{\rmn{spec}}$ values
+close to zero, we are effectively changing the demographics
+of training set by keeping all stars and gradually removing
+galaxies with high redshifts.
 BMC begins to perform worse than morphological separation
@@
 
-The combination technique described in this paper
-can be easily applied to other surveys.
+The problem of star-galaxy classification is a rich area
+for future research.
+It is unclear if sufficient training data will be available in future
+ground-based surveys. Furthermore, in large sky surveys such as DES and LSST,
+photometric quality is not uniform across the sky,
+and a purely morphological classifier alone will not be sufficient,
+especially at faint magnitudes.
 Given the efficacy of our approach,
@@
 
+The authors thank the referee for a careful reading of the manuscript
+and comments that improved this work.
 We thank Ignacio Sevilla for helpful and insightful conversations.
